### 2017-05-20 pds/Paper/paper_pds.tex
initial commit 2

\documentclass{article}
\title{Philosophy of Distributional Semantics\\
\large Best Practices for the Application of Computational Methods to Questions of Meaning}
\date{}
\author{Silvan Hungerb{\"u}hler}
\usepackage{enumitem}
\usepackage{verbatim}
\usepackage{hyperref}
\usepackage{comment}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath}

%\usepackage{biblatex}
%\addbibresource{references_wiggy.bib}
\usepackage{csquotes}
\begin{document}
\maketitle
\begin{abstract}
The increase in availability of data and advances in statistical methods has led to the use Machine Learning (ML) technology in an immense variety of societal domains. ML is used to inform decisions as diverse and life changing as whether somebody obtains financial credit, gets admitted to university, is medically treated or left on probation.
The use of this technology is not always unproblematic, as its careless application can cause serious harm, even if such effects may occur unintentionally.
The seriousness of the potential to systematically disadvantage people has led to a recent debate concerning ethical guidelines and best practices for the \emph{general} use of ML.
Little has been said, however, about the proper use of ML techniques in the specific field of Distributional Semantics (DS). Some researchers have called attention to possible dangers of using ML in DS, as it harbours the potential to reinforce harmful biases along dimensions of race and gender.
The present paper contains an overview over the current state of the \emph{general} discussion 
and draws on these existing insights from different domains of ML to provide guidelines and best practices for computational methods in DS.
As the domain of DS differs in important aspects from other domains of ML, 
we assess these difference with respect on the impact they have on the applicability of general purpose ML guidelines to DS.
\end{abstract}
\section{Introduction}
\paragraph{Motivation}
1-2 paragraphs, talk about application, overal field
\paragraph{Contribution}
largest part of intro, preview of paper, methodology!
\paragraph{Paper Overview}
In \hyperlink{sec2}{Section 2} I roughly situate the use Wittgenstein makes of this philosophical device in the context of his wider philosophy. As this requires a broader view of his philosophy, I rely on two contrasting exegetical efforts.
\hyperlink{sec3}{Section 3} is concerned with tracing Wittgenstein's \textit{Remarks on Frazer's ``Golden Bough''} and their relation to the findings of the previous section. Can Wittgenstein’s views on the proper place of anthropological description as well as his own supposed ethnographic investigations be accommodated within his wider philosophical views? Finally, \hyperlink{sec4}{Section 4} concludes and \hyperlink{sec5}{Section 5} contains the bibliography.

\section{Guidelines for Machine Learning}%\hypertarget{sec2}

\section{Application to DS}%\hypertarget{sec4}
\paragraph{•}
\paragraph{Double-nature of textual corpora}
\section{Related Work}%\hypertarget{sec3}
\section{Conclusion}

\section{References}%\hypertarget{sec5}
%\printbibliography

\end{document}

### 2017-05-20 pds/Paper/paper_pds.tex
added some intro

@@
 %\usepackage{biblatex}
-%\addbibresource{references_wiggy.bib}
+%\addbibresource{pds_references.bib}
 \usepackage{csquotes}
@@
 The increase in availability of data and advances in statistical methods has led to the use Machine Learning (ML) technology in an immense variety of societal domains.
-ML is used to inform decisions as diverse and life changing as whether somebody obtains financial credit, gets admitted to university, is medically treated or left on probation.
+ML is used to inform decisions as diverse and life changing as whether somebody obtains financial credit, is incarcerated, gets admitted to university, is medically treated or legally monitored by the NSA.
 The use of this technology is not always unproblematic, as its careless application can cause serious harm, even if such effects may occur unintentionally.
@@
 \paragraph{Motivation}
-1-2 paragraphs, talk about application, overal field
+1-2 paragraphs, talk about application, overall field
+
+Algorithms are used to aid decisions and take decisions as diverse as the following: %source politicalgrill
+The theory behind these algorithms is \emph{Machine Learning} (ML).\\
+
+Techniques used for the computer-supported study of language meaning, as in \emph{Distributional Semantics}, are a sub-field of ML.
+What does DS diddelydo?
+
+\begin{quote}
+The term “disparate impact” was first used in the 1971 Supreme Court case Griggs v.
+Duke Power Company.
+The Court ruled that, under Title VII of the Civil Rights Act, it was illegal for the company to use intelligence test scores and high school diplomas—factors which were shown to disproportionately favor white applicants and substantially disqualify people of color—to make hiring or promotion decisions, whether or not the company intended the tests to discriminate.
+A key aspect of the Griggs decision was that the power company couldn’t prove their intelligence tests or diploma requirements were actually relevant to the jobs they were hiring for.
+
+In the years since, several disparate impact cases have made their way to the Supreme Court and lower courts, most having to do with employment discrimination.
+This June, the Supreme Court’s decision in Texas Dept.
+of Housing and Community Affairs v.
+Inclusive Communities Project, Inc.
+affirmed the use of the disparate impact theory to fight housing discrimination.
+The Inclusive Communities Project had used a statistical analysis of housing patterns to show that a tax credit program effectively segregated Texans by race.
+Sorelle Friedler, a computer science
+researcher at Haverford College and a fellow at Data \& Society, called the Court’s decision “huge,” both “in favor of civil rights…and in favor of statistics.”
+\end{quote}
+
 \paragraph{Contribution}
 largest part of intro, preview of paper, methodology!
+
+The realization that algorithms fed on big data have the potential to amplify structural discrimination and systematically commit grave errors has led to the emergence of a variety of principles, concrete guidelines and public as well as institutional scrutiny.
+In the present paper we want to draw on the existing efforts made by experts in the fields of ethics, computer science, public policy and law to extend these principles and guidelines to the field of DS.
+Specifically, we will use the \emph{Principles for Accountable Algorithms and a Social Impact Statement for Algorithms} issued by FAT/ML (Fairness, Accountability and Transparency in Machine Learning) - a consortium of researches and experts from private and public organizations working on ML.
+The principles comprise five guiding principles: responsibility, explainability, accuracy, auditibility and fairness. %cite{principles}
+Second, 
+We will assess potential violation of techniques used in DS for each of these principles and, where applicable, issue guidelines to prevent this.
 \paragraph{Paper Overview}
-In \hyperlink{sec2}{Section 2} I roughly situate the use Wittgenstein makes of this philosophical device in the context of his wider philosophy.
-As this requires a broader view of his philosophy, I rely on two contrasting exegetical efforts.
-\hyperlink{sec3}{Section 3} is concerned with tracing Wittgenstein's \textit{Remarks on Frazer's ``Golden Bough''} and their relation to the findings of the previous section.
-Can Wittgenstein’s views on the proper place of anthropological description as well as his own supposed ethnographic investigations be accommodated within his wider philosophical views?
-Finally, \hyperlink{sec4}{Section 4} concludes and \hyperlink{sec5}{Section 5} contains the bibliography.
+The paper is structured as follows: In \hyperlink{sec2}{Section 2} we first give a very brief introduction to the field of ML in general, before providing some background on the particular algorithmic techniques employed in DS.
+Especially for the latter we want to provide overview over societal domains where such technology is currently applied and could conceivably be applied in the future.
+\hyperlink{sec3}{Section 3} contains a discussion of the important ways in which DS, as a particular application of big data analysis, differs from other instances of ML.
+For the present paper these particularities are important insofar as they entail differences when it comes to its risks and accountabilities across the spectrum of ML.
+In \hyperlink{sec4}{Section 4}, then, we try to apply the \emph{Principles for Accountable Algorithms and a Social Impact Statement for Algorithms} by FAT/ML to the specific case of DS.
+\hyperlink{sec5}{Section 5} contains concluding remarks while \hyperlink{sec6}{Section 6} lists the references.
 
@@
 \section{Application to DS}%\hypertarget{sec4}
-\paragraph{•}
-\paragraph{Double-nature of textual corpora}
+\paragraph{What's the same: --}
+\paragraph{What's different: Double-nature of textual corpora}
 \section{Related Work}%\hypertarget{sec3}

### 2017-05-21 pds/Paper/paper_pds.tex
added some structure
updated references. added some ml description

@@
 \begin{abstract}
-The increase in availability of data and advances in statistical methods has led to the use Machine Learning (ML) technology in an immense variety of societal domains.
-ML is used to inform decisions as diverse and life changing as whether somebody obtains financial credit, is incarcerated, gets admitted to university, is medically treated or legally monitored by the NSA.
-The use of this technology is not always unproblematic, as its careless application can cause serious harm, even if such effects may occur unintentionally.
-The seriousness of the potential to systematically disadvantage people has led to a recent debate concerning ethical guidelines and best practices for the \emph{general} use of ML.
-Little has been said, however, about the proper use of ML techniques in the specific field of Distributional Semantics (DS).
+The increase in availability of data and advances in statistical methods has led to the use of Machine Learning (ML) technology in an immense variety of societal domains.
+ML algorithms are used to inform decisions as diverse and life-changing as access to financial credit, incarceration, university admission, medical treatment or the boundaries of espionage by the NSA.
+The use of this technology is not always unproblematic, as its careless application can cause serious harm, even if such effects occur unintentionally.
+The seriousness of the potential to systematically disadvantage people has led to a recent debate concerning ethical guidelines and best practices for the general use of ML.
+Little has been said, however, about the proper use of algorithms in the specific field of Distributional Semantics (DS).
 Some researchers have called attention to possible dangers of using ML in DS, as it harbours the potential to reinforce harmful biases along dimensions of race and gender.
@@
 
-Algorithms are used to aid decisions and take decisions as diverse as the following: %source politicalgrill
+Machine Learning (ML) algorithms are used to aid decisions and take decisions as diverse as the following: %source politicalgrill
 The theory behind these algorithms is \emph{Machine Learning} (ML).\\
@@
 \hyperlink{sec5}{Section 5} contains concluding remarks while \hyperlink{sec6}{Section 6} lists the references.
+\section{Machine Learning and Distributional Semantics}\hypertarget{sec3}{ }
+The term \emph{Machine Learning} is an umbrella term for a variety of statistical methods that, when presented with a series of examples, are able to \emph{learn} how to able to extract generalizable knowledge contained in the examples.
+What is learned are patterns and relationships within the set of examples - often called the \emph{training data} - and the result is a statistical description - or \emph{model} - of the training data.
+\cite{fayyad2001digital}
+\section{Guidelines for Machine Learning}\hypertarget{sec2}{}
 
-\section{Guidelines for Machine Learning}%\hypertarget{sec2}
-
-\section{Application to DS}%\hypertarget{sec4}
+\section{Application to DS}\hypertarget{sec4}{ }
 \paragraph{What's the same: --}
 \paragraph{What's different: Double-nature of textual corpora}
-\section{Related Work}%\hypertarget{sec3}
 \section{Conclusion}
 
-\section{References}%\hypertarget{sec5}
+\section{References}\hypertarget{sec5}{ }
 %\printbibliography

### 2017-05-21 pds/Paper/pds_references.tex
updated references. added some ml description

%pds references

%Ethics/ Best Practices Papers & Articles
@misc{new2016,
  title = {{Living with the algorithms that run our lives} Kernel Description},
  howpublished = {\url{https://www.newscientist.com/article/mg22530073.500-living-with-the-algorithms-that-run-our-lives/
}},
  note = {Accessed: 2017-05-21}
}
@misc{algorithms2016,
  title = {{No one in control} Kernel Description},
  howpublished = {\url{http://thepoliticalgrill.yuku.com/topic/2328/No-one-in-control-The-algorithms-that-run-our-lives#.WSAL_TclHCI
}},
  note = {Accessed: 2017-05-21}
}
@misc{principles,
  title = {{Principles for accountable algorithms} Kernel Description},
  howpublished = {\url{%ttp://www.fatml.org/resources/principles-for-accountable-algorithms
}},
  note = {Accessed: 2017-05-21}
}

%DS Papers

%General ML Texts
@article{fayyad2001digital,
  title={The digital physics of data mining},
  author={Fayyad, Usama},
  journal={Communications of the ACM},
  volume={44},
  number={3},
  pages={62--62},
  year={2001},
  publisher={Association for Computing Machinery, Inc.}
}
@article{domingos2012few,
  title={A few useful things to know about machine learning},
  author={Domingos, Pedro},
  journal={Communications of the ACM},
  volume={55},
  number={10},
  pages={78--87},
  year={2012},
  publisher={ACM}
}
@inproceedings{feldman2015certifying,
  title={Certifying and removing disparate impact},
  author={Feldman, Michael and Friedler, Sorelle A and Moeller, John and Scheidegger, Carlos and Venkatasubramanian, Suresh},
  booktitle={Proceedings of the 21th ACM SIGKDD International Conference on Knowledge Discovery and Data Mining},
  pages={259--268},
  year={2015},
  organization={ACM}
}
@article{barocas2016big,
  title={Big data's disparate impact},
  author={Barocas, Solon and Selbst, Andrew D},
  year={2016}
}

### 2017-05-21 pds/Paper/paper_pds.tex
more intro

@@
 
-Machine Learning (ML) algorithms are used to aid decisions and take decisions as diverse as the following: %source politicalgrill
-The theory behind these algorithms is \emph{Machine Learning} (ML).\\
+As vast amounts of digital data have become available over past decades, algorithms are employed to use the data in order to aid decisions all across society. 
+Algorithms have a say in determining who is granted mortgages, who is released on bail or invited for job applications. 
+\cite{algorithms2016}
+The theory behind these algorithms is \emph{machine learning}.
+Although it seems clear that most designers of machine learning algorithms do not harbor malicious intent, the risk of unwillingly harming people is real.
+Data processed by machine learning techniques often contains societal prejudice and bias, so that unquestioningly using them for decision making can severely reinforce society's existing biases.
+A related issue is that data available to the algorithms may systematically omit information relevant to what it is trying to uncover, as when minority groups are not adequately represented in the data.
+Yet another issue is that machine learning techniques may use useful patterns and relationships in the data that are really mirrorring historical factors of discrimination and marginalization.
+\cite{barocas2016big}
 
-Techniques used for the computer-supported study of language meaning, as in \emph{Distributional Semantics}, are a sub-field of ML.
+These issues have led to an increasing awareness of the societal difficulties of mindlessly employing machine learning techniques.
+
 What does DS diddelydo?
+Recently, machine learning techniques used for the computer-based study of language meaning, as in \emph{Distributional Semantics}, has come under scrutiny.
+Various studies have shown that these methods yield results which are in important ways biased against particular groups as, for example, blacks or women.
 
-\begin{quote}
-The term “disparate impact” was first used in the 1971 Supreme Court case Griggs v.
-Duke Power Company.
-The Court ruled that, under Title VII of the Civil Rights Act, it was illegal for the company to use intelligence test scores and high school diplomas—factors which were shown to disproportionately favor white applicants and substantially disqualify people of color—to make hiring or promotion decisions, whether or not the company intended the tests to discriminate.
-A key aspect of the Griggs decision was that the power company couldn’t prove their intelligence tests or diploma requirements were actually relevant to the jobs they were hiring for.
-
-In the years since, several disparate impact cases have made their way to the Supreme Court and lower courts, most having to do with employment discrimination.
+%\begin{quote}
+%In the years since, several disparate impact cases have made their way to the Supreme Court and lower courts, most having to do with employment discrimination.
 This June, the Supreme Court’s decision in Texas Dept.
@@
 researcher at Haverford College and a fellow at Data \& Society, called the Court’s decision “huge,” both “in favor of civil rights…and in favor of statistics.”
-\end{quote}
+%\end{quote}
 
@@
 \section{Machine Learning and Distributional Semantics}\hypertarget{sec3}{ }
-The term \emph{Machine Learning} is an umbrella term for a variety of statistical methods that, when presented with a series of examples, are able to \emph{learn} how to able to extract generalizable knowledge contained in the examples.
-What is learned are patterns and relationships within the set of examples - often called the \emph{training data} - and the result is a statistical description - or \emph{model} - of the training data.
-\cite{fayyad2001digital}
+The term \emph{Machine Learning} is an umbrella term for a variety of statistical methods that, when presented with a series of examples, are able to \emph{learn} from them.
+That is, the the set of examples - often called the \emph{training data} - is not simply memorized but generalizable knowledge extracted from it.
+\cite{domingos2012few}
+ What is learned are patterns and relationships within the training data and the result is a statistical description - or \emph{model} - of the training data. \cite{fayyad2001digital}  
+ The
 \section{Guidelines for Machine Learning}\hypertarget{sec2}{}

### 2017-05-21 pds/Paper/paper_pds.tex
final commit of the day. sort of an intro

@@
 
-These issues have led to an increasing awareness of the societal difficulties of mindlessly employing machine learning techniques.
-
-What does DS diddelydo?
-Recently, machine learning techniques used for the computer-based study of language meaning, as in \emph{Distributional Semantics}, has come under scrutiny.
-Various studies have shown that these methods yield results which are in important ways biased against particular groups as, for example, blacks or women.
-
+Recently, machine learning techniques used for the computer-based study of language meaning - a field known as \emph{Computational Distributional Semantics} - has come under scrutiny.
+Various studies have shown that these methods yield results which are in important ways biased against particular groups as, for example, women. \cite{google}
 %\begin{quote}
@@
 Sorelle Friedler, a computer science
-researcher at Haverford College and a fellow at Data \& Society, called the Court’s decision “huge,” both “in favor of civil rights…and in favor of statistics.”
+%researcher at Haverford College and a fellow at Data \& Society, called the Court’s decision “huge,” both “in favor of civil rights…and in favor of statistics.”
 %\end{quote}
 
-\paragraph{Contribution}
-largest part of intro, preview of paper, methodology!
+%\paragraph{Contribution}
+%largest part of intro, preview of paper, methodology!
 
@@
 Specifically, we will use the \emph{Principles for Accountable Algorithms and a Social Impact Statement for Algorithms} issued by FAT/ML (Fairness, Accountability and Transparency in Machine Learning) - a consortium of researches and experts from private and public organizations working on ML.
-The principles comprise five guiding principles: responsibility, explainability, accuracy, auditibility and fairness. %cite{principles}
-Second, 
-We will assess potential violation of techniques used in DS for each of these principles and, where applicable, issue guidelines to prevent this.
-\paragraph{Paper Overview}
+The principles comprise five guiding principles: responsibility, explainability, accuracy, auditibility and fairness \cite{principles}.
+We will assess potential violation of techniques used in distributional semantics for each of these principles and, where applicable, issue guidelines to prevent their negative impact upon vulnerable groups.
+%\paragraph{Paper Overview}
+
 The paper is structured as follows: In \hyperlink{sec2}{Section 2} we first give a very brief introduction to the field of ML in general, before providing some background on the particular algorithmic techniques employed in DS.

### 2017-05-23 pds/Paper/pds_references.tex
final commit of the day. sort of an intro
added some flesh to that there bone

@@
 
+@misc{principles,
+  title = {{Discrimination algorithms disparate impact} Kernel Description},
+  howpublished = {\url{https://www.theatlantic.com/business/archive/2015/09/discrimination-algorithms-disparate-impact/403969/
+}},
+  note = {Accessed: 2017-05-22}
+}
+
 %DS Papers
+@article{wagner2015s,
+  title={It's a man's wikipedia? Assessing gender inequality in an online encyclopedia},
+  author={Wagner, Claudia and Garcia, David and Jadidi, Mohsen and Strohmaier, Markus},
+  journal={arXiv preprint arXiv:1501.06307},
+  year={2015}
+}
+@article{harris1954distributional,
+  title={Distributional structure},
+  author={Harris, Zellig S},
+  journal={Word},
+  volume={10},
+  number={2-3},
+  pages={146--162},
+  year={1954},
+  publisher={Taylor \& Francis}
+}
+@article{boleda2016formal,
+  title={Formal Distributional Semantics: Introduction to the Special Issue},
+  author={Boleda, Gemma and Herbelot, Aur{\'e}lie},
+  journal={Computational Linguistics},
+  year={2016},
+  publisher={MIT Press}
+}
 

### 2017-05-27 pds/Paper/paper_pds.tex
some updates

@@
 \section{Introduction}
-\paragraph{Motivation}
-1-2 paragraphs, talk about application, overall field
-
-As vast amounts of digital data have become available over past decades, algorithms are employed to use the data in order to aid decisions all across society. 
+%1-2 paragraphs, talk about application, overall field
+\paragraph{Place of Machine Learning in society}
+As vast amounts of digital data have become available over past decades, algorithms are employed to use the data in order to take and inform decisions all across society. 
 Algorithms have a say in determining who is granted mortgages, who is released on bail or invited for job applications. 
@@
 The theory behind these algorithms is \emph{machine learning}.
-Although it seems clear that most designers of machine learning algorithms do not harbor malicious intent, the risk of unwillingly harming people is real.
+Although most designers of machine learning algorithms do not harbor malicious intent, the risk of unwillingly harming people is real.
+\paragraph{Harmful Effects}
+-Privacy - where does data get taken from, who gets to see it, treatment\\
+-Harming Workers - effects of automation\\
+-Questionable Use - military/espionage application\\
+-Reeinforcing discrimination and bias
+
+have the potential to amplify structural discrimination and systematically commit grave errors
+
 Data processed by machine learning techniques often contains societal prejudice and bias, so that unquestioningly using them for decision making can severely reinforce society's existing biases.
@@
 
-Recently, machine learning techniques used for the computer-based study of language meaning - a field known as \emph{Computational Distributional Semantics} - has come under scrutiny.
-Various studies have shown that these methods yield results which are in important ways biased against particular groups as, for example, women. \cite{google}
+\paragraph{2b moved}
+Recently, machine learning techniques used for the computer-based study of language meaning - a field known as \emph{computational distributional semantics} - has come under scrutiny.
+Various studies have shown that these methods yield results which are in important ways biased against particular groups as, for example, women \cite{google} \cite{wagner2015s}.
 %\begin{quote}
@@
 %largest part of intro, preview of paper, methodology!
+\paragraph{Public Attention is in ML minor extent in NLP, not in DS}
+The realization that algorithms fed on big data come with a host of problematic issues has led to the emergence of a variety of principles, concrete guidelines and public as well as institutional scrutiny.
 
-The realization that algorithms fed on big data have the potential to amplify structural discrimination and systematically commit grave errors has led to the emergence of a variety of principles, concrete guidelines and public as well as institutional scrutiny.
+\paragraph{Contribution: Drawing on ML discussion for DS, Methodology}
 In the present paper we want to draw on the existing efforts made by experts in the fields of ethics, computer science, public policy and law to extend these principles and guidelines to the field of DS.
-Specifically, we will use the \emph{Principles for Accountable Algorithms and a Social Impact Statement for Algorithms} issued by FAT/ML (Fairness, Accountability and Transparency in Machine Learning) - a consortium of researches and experts from private and public organizations working on ML.
-The principles comprise five guiding principles: responsibility, explainability, accuracy, auditibility and fairness \cite{principles}.
+Specifically, we will use the \emph{Principles for Accountable Algorithms and a Social Impact Statement for Algorithms} (PAASISA) issued by an organization called \emph{Fairness, Accountability and Transparency in Machine Learning} - a consortium of researches and experts from private and public organizations working on machine learning.
+The PAASISA comprise five guiding principles: responsibility, explainability, accuracy, auditibility and fairness \cite{principles}.
 We will assess potential violation of techniques used in distributional semantics for each of these principles and, where applicable, issue guidelines to prevent their negative impact upon vulnerable groups.
-%\paragraph{Paper Overview}
 
-The paper is structured as follows: In \hyperlink{sec2}{Section 2} we first give a very brief introduction to the field of ML in general, before providing some background on the particular algorithmic techniques employed in DS.
+\paragraph{Paper Overview}
+The paper is structured as follows: In \hyperlink{sec2}{Section 2} we first give a very brief introduction to the field of machine learning in general, before providing some background on the algorithmic techniques employed in distributional semantics.
 Especially for the latter we want to provide overview over societal domains where such technology is currently applied and could conceivably be applied in the future.
@@
 \section{Machine Learning and Distributional Semantics}\hypertarget{sec3}{ }
-The term \emph{Machine Learning} is an umbrella term for a variety of statistical methods that, when presented with a series of examples, are able to \emph{learn} from them.
-That is, the the set of examples - often called the \emph{training data} - is not simply memorized but generalizable knowledge extracted from it.
-\cite{domingos2012few}
- What is learned are patterns and relationships within the training data and the result is a statistical description - or \emph{model} - of the training data. \cite{fayyad2001digital}  
- The
+The term \emph{Machine Learning} is an umbrella term for a variety of statistical methods that are able to \emph{learn} from a series of examples presented to them.
+The the set of examples - often called the \emph{training data} - is not simply memorized but the machine learning algorithm extracts generalizable knowledge \cite{domingos2012few}.
+ What the algorithm learns are patterns and relationships within the training data and the result of the learning process is a statistical description - or \emph{model} - of the training data \cite{fayyad2001digital}.
+
+ \paragraph{Distributional Semantics}
+Distributional semantics is a linguistic theory of meaning looking to represent lexical meaning of expressions as a function of the context in which they appear.
+It theoretically foundation is the so-called \emph{distributional hypothesis} which holds that two expressions are similar in meaning, if they occur in similar contexts \cite{harris1954distributional}.
+Althought it was originally developed as a distributional theory - in the sense that ...
+- in the middle of the 20th century, it can be seen as following the footsteps of structural linguists, notably Bloomfield \cite{bloomfield}
+
+More info on DS. use \cite{boleda2016formal}
+
+The theoretical framework of distributional semantics has become widely used and further developed as the advent and success of computational methods have enabled linguist to build large-scale models of linguistic meaning using digitalised corpora of text.
+
+\paragraph{ML methods in DS, some examples}
+
+\paragraph{Where models of meaning could have an import}
+
 \section{Guidelines for Machine Learning}\hypertarget{sec2}{}
+Algorithmically informed decisions are used in ever more domains of the private and public sectors and thus have the potential to impact society in significant ways.
+In many cases of undesirable effects that come with these technologies, legislation is behind in adequately protecting people negatively affected by it.
+Often a clear legal base is lacking to give clear guidelines.
+
+Given this state of affairs, it is important to hold the people who develop or implement such algorithmic systems accountable to the public. The FAT-ML concisely write in their PAA \cite{principles}
+\begin{quote}
+Algorithms and the data that drive them are designed and created by people -- There is always a human ultimately responsible for decisions made or informed by an algorithm.
+''The algorithm did it'' is not an acceptable excuse if algorithmic systems make mistakes or have undesired consequences, including from machine-learning processes.
+\end{quote}
+
+The document then goes on to name five principles that should guide developders of algorithms in designing and implementing their systems.
+The authors of PAA urge creators to issue a Social Impact Statement wherein they detail the effect of their product has on society with respect to the five principles.
+In this way the principles are associeated with best practices in the ...
+
+We will now briefly go over the five principles to explaint their content.
 

### 2017-05-28 pds/Paper/paper_pds.tex
added some intro

@@
 The use of this technology is not always unproblematic, as its careless application can cause serious harm, even if such effects occur unintentionally.
-The seriousness of the potential to systematically disadvantage people has led to a recent debate concerning ethical guidelines and best practices for the general use of ML.
+The seriousness of the potential to disrespect people's privacy and integrity or systematically disadvantage specific groups has led to a recent debate concerning ethical guidelines and best practices for the general use of ML.
 Little has been said, however, about the proper use of algorithms in the specific field of Distributional Semantics (DS).
 Some researchers have called attention to possible dangers of using ML in DS, as it harbours the potential to reinforce harmful biases along dimensions of race and gender.
-The present paper contains an overview over the current state of the \emph{general} discussion 
-and draws on these existing insights from different domains of ML to provide guidelines and best practices for computational methods in DS.
-As the domain of DS differs in important aspects from other domains of ML, 
+The present paper wants to point to one particular attempt to provide general-purpose guidelines for ML and draw on these existing insights to suggest guidelines and best practices for computational methods in DS.
+As the domain of DS differs in important aspects from other domains of ML, notably Natural Language Processing (NLP),
 we assess these difference with respect on the impact they have on the applicability of general purpose ML guidelines to DS.
@@
 The theory behind these algorithms is \emph{machine learning}.
-Although most designers of machine learning algorithms do not harbor malicious intent, the risk of unwillingly harming people is real.
+Although most inventors and designers of machine learning algorithms do not harbor malicious intent, the risk of unwillingly harming people is real.
 \paragraph{Harmful Effects}
--Privacy - where does data get taken from, who gets to see it, treatment\\
--Harming Workers - effects of automation\\
--Questionable Use - military/espionage application\\
+There are several different sources of harmful effects that these techniques can bring about.
+
+A first concern is the issue of privacy.
+Since very large amounts of data are required for these methods to be effective, great efforts are devoted to collecting large data sets.
+Depending on the issue at hand, these data sets may contain highly sensitive information about individuals or organizations.
+It is thus of importance who gets to see these data and in what they are protected from third parties accessing them.
+There have been reported cases of researchers working with large corpora of texts containing peoples' complete medical record who sent these files indiscriminately around by email.
+A second aspect of privacy is how the data is collected in the first place.
+It is a question of ethical concern whether the individuals whose data is collected agree to their personal information being used.
+What the standards should be here isstill  far from clear as automated data-mining techniques gain popularity.
+
+A second point concerns the disruptive potential machine learning techniques have on labor markets.
+As ever more cognitive tasks can be automated, the economic landscape of service sector economies is bound undergo change.
+Some workers will without a doubt be disadvantaged by these changes.
+It must concern the inventors and implentors of such technology what happens here.
+
+A third difficult point is the so-called \emph{double use} problem.
+New technology's use is often underspecified and could be used for benefit of humanity or to its detriment.
+Chat-bots developed by researchers in automated natural language processing, to name a concrete example, can be used to assist people to access governmental health-care serices, yet they have reportedly been used by the Mexican government to disrupt political discussions deemed subversive on social media.
+\cite{•}
 -Reeinforcing discrimination and bias
@@
 
-\paragraph{Contribution: Drawing on ML discussion for DS, Methodology}
+\paragraph{Contribution: Drawing on ML discussion for DS}
 In the present paper we want to draw on the existing efforts made by experts in the fields of ethics, computer science, public policy and law to extend these principles and guidelines to the field of DS.
@@
 We will assess potential violation of techniques used in distributional semantics for each of these principles and, where applicable, issue guidelines to prevent their negative impact upon vulnerable groups.
+
+We want to focus on aspects of discrimination and bias in the outcome because this seems least explored.
+Issues concerning privacy etc.
+seem to follow the same lines as general purpose ML, or even research, ethical principles while the issue of discrimination and bias has a special relation to DS.
+
+\paragraph{Methodoology}
 
@@
 \hyperlink{sec5}{Section 5} contains concluding remarks while \hyperlink{sec6}{Section 6} lists the references.
-\section{Machine Learning and Distributional Semantics}\hypertarget{sec3}{ }
+\section{Machine Learning, Natural Language Processing and Distributional Semantics}\hypertarget{sec3}{ }
+\paragraph{What is ML?}
 The term \emph{Machine Learning} is an umbrella term for a variety of statistical methods that are able to \emph{learn} from a series of examples presented to them.
@@
  What the algorithm learns are patterns and relationships within the training data and the result of the learning process is a statistical description - or \emph{model} - of the training data \cite{fayyad2001digital}.
-
- \paragraph{Distributional Semantics}
+\paragraph{What is NLP?}
+\paragraph{Relation: ML - NLP}
+\paragraph{What is Distributional Semantics?}
 Distributional semantics is a linguistic theory of meaning looking to represent lexical meaning of expressions as a function of the context in which they appear.
@@
 - in the middle of the 20th century, it can be seen as following the footsteps of structural linguists, notably Bloomfield \cite{bloomfield}
-
 More info on DS. use \cite{boleda2016formal}
@@
 
-\paragraph{ML methods in DS, some examples}
+\paragraph{Relation: NLP - DS}
+DS is naturally concerned with a narrower topic: meaning.\\
+Meaning is a notoriously difficult topic conceptually.\\
+DS sits at the intersection of ethically touchy topics: capturing/reporting/describing bias leads to reenforcing bias.
+results can be highly sensitive to data selection.
+even if results are valid there are issues of over/under-representation (scottish accent),
 
-\paragraph{Where models of meaning could have an import}
+\paragraph{The state of the ethical discussion in ML, NLP and DS}
+-ML well underway. series of legal, political and cs studies on the topic. many articles in non-technical or academic media. be specific, mention FATML. use nlp ethic article to get overview.\\
+-NLP has seen some movement in the last couple of years. some conferences on ethical issues. even concrete proposals for best practices.\\
+-DS in particular has seen little attention.
 
+\paragraph{What good should come from looking at DS?}
+-focus not on dual-use or bad treatment of mechanical turks, but the question how bias in meaning is dangerous.\\
+-Why not ethics by design
 \section{Guidelines for Machine Learning}\hypertarget{sec2}{}
+\paragraph{Introduce FATML}
 Algorithmically informed decisions are used in ever more domains of the private and public sectors and thus have the potential to impact society in significant ways.
 In many cases of undesirable effects that come with these technologies, legislation is behind in adequately protecting people negatively affected by it.
-Often a clear legal base is lacking to give clear guidelines.
 
@@
 The authors of PAA urge creators to issue a Social Impact Statement wherein they detail the effect of their product has on society with respect to the five principles.
-In this way the principles are associeated with best practices in the ...
+In this way the principles are associated with best practices in the ...
+
+\paragraph{5 Principles, quick recap of each of them}
 

### 2017-05-28 pds/Paper/paper_pds.tex
intro in decent shape

@@
 Chat-bots developed by researchers in automated natural language processing, to name a concrete example, can be used to assist people to access governmental health-care serices, yet they have reportedly been used by the Mexican government to disrupt political discussions deemed subversive on social media.
-\cite{•}
--Reeinforcing discrimination and bias
+\cite{leidner2017ethical}
 
-have the potential to amplify structural discrimination and systematically commit grave errors
-
+Finally, machine learning harbours the potential to amplify structural discrimination and systematically disadvantage people because of their race, gender or other social category.
+If the models which are trained algorithmically are fed biased data, then predictions or classifications they make will reflect that bias.
 Data processed by machine learning techniques often contains societal prejudice and bias, so that unquestioningly using them for decision making can severely reinforce society's existing biases.
@@
 
-\paragraph{2b moved}
-Recently, machine learning techniques used for the computer-based study of language meaning - a field known as \emph{computational distributional semantics} - has come under scrutiny.
-Various studies have shown that these methods yield results which are in important ways biased against particular groups as, for example, women \cite{google} \cite{wagner2015s}.
+In the present essay, we will mostly be concerned with the last of these problematic issues surrounding machine learning.
+
 %\begin{quote}
@@
 \paragraph{Public Attention is in ML minor extent in NLP, not in DS}
-The realization that algorithms fed on big data come with a host of problematic issues has led to the emergence of a variety of principles, concrete guidelines and public as well as institutional scrutiny.
+Many of the issues lately mentioned are of great importance to modern society at large.
+Consequently, with the rise of big data people working at the intersection of information technology and ethics have attempted to lay down foundations on how to handle these technologies.
+Many of these efforts have naturally been directed at the broad topic of machine learning in general.\cite{leidner2017ehtical} Only very recently the discussion has turned on machine learning used for analysis and modeling of language - often referred to as \emph{natural language processing}.
+Presumably because this is only one of many particular fields where machine learning is applied and the discussion of ethics natural language processing is bound pertain to less cases of applications throughout society, it took some time for the discussion to reach this field.
+
+Recently, however, a discussion has opened up in a hitherto completely unmentioned domain.
+Machine learning techniques used for the computer-based study of language meaning - a field known as \emph{computational distributional semantics} - has come under scrutiny.
+Various studies have shown that these methods yield results which are in important ways biased against particular groups as, for example, women \cite{google} \cite{wagner2015s}.The realization that algorithms fed on big data come with a host of problematic issues has led to the emergence of a variety of principles, concrete guidelines and public as well as institutional scrutiny.
 
@@
 
-We want to focus on aspects of discrimination and bias in the outcome because this seems least explored.
-Issues concerning privacy etc.
-seem to follow the same lines as general purpose ML, or even research, ethical principles while the issue of discrimination and bias has a special relation to DS.
+We want to focus on aspects of discrimination and bias in the outcome because these issues are least explored.
+Debates concerning privacy etc.
+would seem to follow the same lines as the discussion to establish general purpose ethical principles for machine learning, while the issue of discrimination and bias has a special relation to DS.
 
 \paragraph{Methodoology}
+The method employed by the present contribution is one of scanning previous efforts in a similar vein, identifying useful parts and applying them to ethical questions surrounding distributional semantics.
+There will be no foundational efforts undertaken to sustain the validity of the principles thus transferred.
+Their validity is presupposed.
+The analysis contained in this contribution merely serves to disucss the usefulness of the principles in particular domain, and assert it where due.
 

### 2017-05-28 pds/Paper/paper_pds.tex
advanced some in second part

@@
 \paragraph{Contribution: Drawing on ML discussion for DS}
-In the present paper we want to draw on the existing efforts made by experts in the fields of ethics, computer science, public policy and law to extend these principles and guidelines to the field of DS.
-Specifically, we will use the \emph{Principles for Accountable Algorithms and a Social Impact Statement for Algorithms} (PAASISA) issued by an organization called \emph{Fairness, Accountability and Transparency in Machine Learning} - a consortium of researches and experts from private and public organizations working on machine learning.
+In the present paper we want to draw on the existing efforts made by experts in the fields of ethics, computer science, public policy and law to extend these principles and guidelines to the field of distributional semantics.
+Specifically, we will use the \emph{Principles for Accountable Algorithms and a Social Impact Statement for Algorithms} (PAASISA) issued by the organization \emph{Fairness, Accountability and Transparency in Machine Learning} (FATML) - a consortium of researches and experts from private and public organizations working on machine learning.
 The PAASISA comprise five guiding principles: responsibility, explainability, accuracy, auditibility and fairness \cite{principles}.
@@
 
-We want to focus on aspects of discrimination and bias in the outcome because these issues are least explored.
+Currently, we want to focus on aspects of discrimination and bias in the outcome of distributional semantics models for two reasons.
+For one, these are the least explored issues for distributional semantics.
+Many of the ideas, guidelines and principles concerning privacy, double use or economic impact are directly transferable from the general machine learning debate to any sub-field that similarly employs big data and algorithms, as does computational distributional semantics.
 Debates concerning privacy etc.
-would seem to follow the same lines as the discussion to establish general purpose ethical principles for machine learning, while the issue of discrimination and bias has a special relation to DS.
+can follow the same lines as the discussion to establish general purpose ethical principles for machine learning.
+The issue of discrimination and bias, however, has a special relation to distributional semantics because it occupies in a unique position of computational language research.
+The central concern of distributional semantics, as its name gives away, is the concept of meaning.
+As we will argue presently, this relates it in a special way to many sensitive topics in the general machine learning discussion.
+Put succinctly, it sits at the intersection of concerns pertaining to data selection, validity or questionability of computational methods employed, interpretation of results and usage thereof.
 
@@
 Their validity is presupposed.
-The analysis contained in this contribution merely serves to disucss the usefulness of the principles in particular domain, and assert it where due.
+The analysis contained in this contribution merely serves to discuss the usefulness of the principles in particular domain, and assert it where due.
 
 \paragraph{Paper Overview}
-The paper is structured as follows: In \hyperlink{sec2}{Section 2} we first give a very brief introduction to the field of machine learning in general, before providing some background on the algorithmic techniques employed in distributional semantics.
-Especially for the latter we want to provide overview over societal domains where such technology is currently applied and could conceivably be applied in the future.
+The paper is structured as follows: In \hyperlink{sec2}{Section 2} we first give a very brief description of machine learning, natural language processing and distributional semantics as well as their relation to each other.
+Especially for distributional semantics we want to provide overview over societal domains where such technology is currently applied and could conceivably be applied in the future.
 \hyperlink{sec3}{Section 3} contains a discussion of the important ways in which DS, as a particular application of big data analysis, differs from other instances of ML.
@@
 \paragraph{What is NLP?}
+Natural language processing is a field of research that uses computational tools to study natural language text or speech.
+The knowledge thus obtained can be used to develop computer systems to serve a wide variety of purposes.
+Principal applications include translation between natural languages, artificial intelligence, user interfaces, like chat-bots, for example, summarization of text or speech recognition \cite{chowdhury2003natural}.
 \paragraph{Relation: ML - NLP}
+The area of natural language processing frequently serves itself of methods from machine learning, for example to process very large textual corpora.
+Indeed, parts of natural language processing can be thought of as particular applications of machine learning techniques to the domain of natural language.
 \paragraph{What is Distributional Semantics?}
@@
 It theoretically foundation is the so-called \emph{distributional hypothesis} which holds that two expressions are similar in meaning, if they occur in similar contexts \cite{harris1954distributional}.
-Althought it was originally developed as a distributional theory - in the sense that ...
+%Althought it was originally developed as a distributional theory - in the sense that ...
 - in the middle of the 20th century, it can be seen as following the footsteps of structural linguists, notably Bloomfield \cite{bloomfield}
-More info on DS. use \cite{boleda2016formal}
 
+The theory of distributional semantics has been used successfully in computational semantics, as it has become possible to obtain semantic representations immediately from natural language data .
 The theoretical framework of distributional semantics has become widely used and further developed as the advent and success of computational methods have enabled linguist to build large-scale models of linguistic meaning using digitalised corpora of text.
+Following the distributional hypothesis, the meaning of an expression is a function of the context in which it occurs, where context can be construed in different ways.
+Most typically, it is other words surrounding an expression in a text - the sentence in which it appears, for example.
+Context is not restricted to such simple analysis, however, as it can possibly contain more complex syntactic relations within a text, or even visual, textual or auditory information gathered from data.
+\cite{boleda2016formal}
+Distributional semantics, then, takes the context in which an expression occurs and abstracts away from the particular contexts in a large corpus of natural language data.
+The mathematical vehicle for representing these abstractions are commonly vectors, but there are also other algebraic objects possible, as matrices or tensors.
+The meaning of an expression, as a function of its context, is then represented by the distribution across the dimensions of the vector.
+A set of words thus represented builds a semantic space - mathematically, a vector space - wherein one can study the semantic relationship between the expressions because some semantic relationships between expressions are mirrored by geometric relationship in these spaces.
 
 \paragraph{Relation: NLP - DS}
-DS is naturally concerned with a narrower topic: meaning.\\
-Meaning is a notoriously difficult topic conceptually.\\
+Distributional semantics is concerned with a specific topic in the study of natural language, namely, meaning. It could thus be seen as one of many sub-fields of natural language processing research. 
+It is notoriously difficult to come to grips with the concept of meaning from philosophical standpoint.
 DS sits at the intersection of ethically touchy topics: capturing/reporting/describing bias leads to reenforcing bias.

### 2017-05-28 pds/Paper/pds_references.tex
advanced some in second part

--- 2017-05-21 pds/Paper/pds_references.tex
@@
 
+@misc{principles,
+  title = {{Discrimination algorithms disparate impact} Kernel Description},
+  howpublished = {\url{https://www.theatlantic.com/business/archive/2015/09/discrimination-algorithms-disparate-impact/403969/
+}},
+  note = {Accessed: 2017-05-22}
+}
+@PROCEEDINGS{proceedings,
+  title = {Proc.\ o.t. First Workshop on Ethics in NLP
+      (Valencia, Spain, April 2017)},
+  booktitle = {Proc.\ o.t. First Workshop on Ethics in NLP
+      (Valencia, Spain, April 2017)},
+  volume = 1,
+  series = {},
+  publisher = {},
+  address = {Valencia, Spain},
+  year = {2017}
+}
+
+@inproceedings{leidner2017ethical,
+  author =       {Jochen L. Leidner, Vassilis Plachouras},
+  title =        {Ethical by Design: Ethics Best Practices for Natural Language Processing},
+  crossref =     {proceedings},
+  pages =        {9--18},
+}
+
 %DS Papers
+@article{wagner2015s,
+  title={It's a man's wikipedia? Assessing gender inequality in an online encyclopedia},
+  author={Wagner, Claudia and Garcia, David and Jadidi, Mohsen and Strohmaier, Markus},
+  journal={arXiv preprint arXiv:1501.06307},
+  year={2015}
+}
+@article{harris1954distributional,
+  title={Distributional structure},
+  author={Harris, Zellig S},
+  journal={Word},
+  volume={10},
+  number={2-3},
+  pages={146--162},
+  year={1954},
+  publisher={Taylor \& Francis}
+}
+@article{boleda2016formal,
+  title={Formal Distributional Semantics: Introduction to the Special Issue},
+  author={Boleda, Gemma and Herbelot, Aur{\'e}lie},
+  journal={Computational Linguistics},
+  year={2016},
+  publisher={MIT Press}
+}
+@article{chowdhury2003natural,
+  title={Natural language processing},
+  author={Chowdhury, Gobinda G},
+  journal={Annual review of information science and technology},
+  volume={37},
+  number={1},
+  pages={51--89},
+  year={2003},
+  publisher={Wiley Online Library}
+}
 

### 2017-05-28 pds/Paper/pds_references.tex
good portion of 2 added

--- 2017-05-21 pds/Paper/pds_references.tex
@@
 
+@misc{principles,
+  title = {{Discrimination algorithms disparate impact} Kernel Description},
+  howpublished = {\url{https://www.theatlantic.com/business/archive/2015/09/discrimination-algorithms-disparate-impact/403969/
+}},
+  note = {Accessed: 2017-05-22}
+}
+@article{sweeney2013discrimination,
+  title={Discrimination in online ad delivery},
+  author={Sweeney, Latanya},
+  journal={Queue},
+  volume={11},
+  number={3},
+  pages={10},
+  year={2013},
+  publisher={ACM}
+}
+
+@PROCEEDINGS{proceedings,
+  title = {Proc.\ o.t. First Workshop on Ethics in NLP
+      (Valencia, Spain, April 2017)},
+  booktitle = {Proc.\ o.t. First Workshop on Ethics in NLP
+      (Valencia, Spain, April 2017)},
+  volume = 1,
+  series = {},
+  publisher = {},
+  address = {Valencia, Spain},
+  year = {2017}
+}
+
+@inproceedings{leidner2017ethical,
+  author =       {Jochen L. Leidner, Vassilis Plachouras},
+  title =        {Ethical by Design: Ethics Best Practices for Natural Language Processing},
+  crossref =     {proceedings},
+  pages =        {9--18},
+}
+
 %DS Papers
+@article{wagner2015s,
+  title={It's a man's wikipedia? Assessing gender inequality in an online encyclopedia},
+  author={Wagner, Claudia and Garcia, David and Jadidi, Mohsen and Strohmaier, Markus},
+  journal={arXiv preprint arXiv:1501.06307},
+  year={2015}
+}
+@article{harris1954distributional,
+  title={Distributional structure},
+  author={Harris, Zellig S},
+  journal={Word},
+  volume={10},
+  number={2-3},
+  pages={146--162},
+  year={1954},
+  publisher={Taylor \& Francis}
+}
+@article{boleda2016formal,
+  title={Formal Distributional Semantics: Introduction to the Special Issue},
+  author={Boleda, Gemma and Herbelot, Aur{\'e}lie},
+  journal={Computational Linguistics},
+  year={2016},
+  publisher={MIT Press}
+}
+@article{chowdhury2003natural,
+  title={Natural language processing},
+  author={Chowdhury, Gobinda G},
+  journal={Annual review of information science and technology},
+  volume={37},
+  number={1},
+  pages={51--89},
+  year={2003},
+  publisher={Wiley Online Library}
+}
+@inproceedings{bolukbasi2016man,
+  title={Man is to computer programmer as woman is to homemaker? debiasing word embeddings},
+  author={Bolukbasi, Tolga and Chang, Kai-Wei and Zou, James Y and Saligrama, Venkatesh and Kalai, Adam T},
+  booktitle={Advances in Neural Information Processing Systems},
+  pages={4349--4357},
+  year={2016}
+}
 

### 2017-05-28 pds/Paper/paper_pds.tex
final commit

@@
 Distributional semantics is concerned with a specific topic in the study of natural language, namely, meaning. It could thus be seen as one of many sub-fields of natural language processing research. 
-It is notoriously difficult to come to grips with the concept of meaning from philosophical standpoint.
-DS sits at the intersection of ethically touchy topics: capturing/reporting/describing bias leads to reenforcing bias.
-results can be highly sensitive to data selection.
-even if results are valid there are issues of over/under-representation (scottish accent),
+From a philosophical standpoint, it is notoriously difficult to come to grips with the concept of meaning .
+Distributional semantics sits at the intersection of ethically touchy topics. 
+First off, consider the situation where a meaning space is constructed based on some seemingly innocuous corpus and ends up containing severe bias along lines of gender or racial stereotypes.
+A recent study has found, for example, that a model trained on Google News articles display significant bias along binary gender dimensions \cite{bolukbasi2016man}.
+This raises a host of questions concerning the provenance of the bias, the adequate response by the responsible distributional semanticist and, not to mention, the nature of meaning.
+
+Does one have to live with the fact that the meanings discovered in some corpus systematically contain unfavourable connotations for some group of people?
+After all, the corpus was produced by a particular society, so it should be expected that it mirror that society's prejudices.
+Is the stance "the model picked up what it is supposed to pick up, blame society" a fair one, or has the computational linguist the moral obligation to rid her model of biased meaning representations?
+
+Fair and square, the model contains bias because the corpus contains bias.
+What if the semantic model is further used for some interaction within that same society that taught it - via corpus - what it means to be a \emph{woman}, will the bias be further re-enforced?
+It is has been reported that when one queries Google with the name of a person, then Google is more likely to show advertisement related to that person being arrested if the name's connotation is black \cite{sweeney2013discrimination}.
+%A related issue is that even if results are valid, the of over-, or, respectively, under-representation of specific items in the data can disadvantage group.
 
 \paragraph{The state of the ethical discussion in ML, NLP and DS}
--ML well underway. series of legal, political and cs studies on the topic. many articles in non-technical or academic media. be specific, mention FATML. use nlp ethic article to get overview.\\
--NLP has seen some movement in the last couple of years. some conferences on ethical issues. even concrete proposals for best practices.\\
+For the domain general use of machine learning techniques there exists a series of studies by now from legal, political computer science experts which specify principles and guidelines.
+Furthermore, many non-technical articles have been published outside of academic channels.
+be specific, mention FATML. use nlp ethic article to get overview.\\
+
+In the case of natural language processing there has been  has seen some movement in the last couple of years.
+For machine learning and natural language processing there are several  conferences on ethical issues by now.
+http://www.ethicsinnlp.org/related-conferences
+
+ even concrete proposals for best practices.\\
 -DS in particular has seen little attention.

### 2017-05-29 pds/Paper/paper_pds.tex
get shits saved

@@
 The increase in availability of data and advances in statistical methods has led to the use of Machine Learning (ML) technology in an immense variety of societal domains.
-ML algorithms are used to inform decisions as diverse and life-changing as access to financial credit, incarceration, university admission, medical treatment or the boundaries of espionage by the NSA.
+ML algorithms are used to inform decisions as diverse and life-changing as access to financial credit, incarceration, university admission, medical treatment or the legal boundaries of espionage by the NSA.
 The use of this technology is not always unproblematic, as its careless application can cause serious harm, even if such effects occur unintentionally.
@@
 The present paper wants to point to one particular attempt to provide general-purpose guidelines for ML and draw on these existing insights to suggest guidelines and best practices for computational methods in DS.
-As the domain of DS differs in important aspects from other domains of ML, notably Natural Language Processing (NLP),
-we assess these difference with respect on the impact they have on the applicability of general purpose ML guidelines to DS.
+As the domain of DS differs in important aspects from other applications of ML, notably also for Natural Language Processing (NLP),
+this involves assessing  the applicability of general purpose ML guidelines to DS with these difference in mind.
 \end{abstract}
@@
 \cite{algorithms2016}
-The theory behind these algorithms is \emph{machine learning}.
-Although most inventors and designers of machine learning algorithms do not harbor malicious intent, the risk of unwillingly harming people is real.
+The theory behind many of these algorithms is \emph{machine learning} (ML).
+Although most inventors and designers of ML algorithms certainly do not harbor malicious intent, the risk of unwillingly harming people is real.
+
 \paragraph{Harmful Effects}
 There are several different sources of harmful effects that these techniques can bring about.
-
-A first concern is the issue of privacy.
-Since very large amounts of data are required for these methods to be effective, great efforts are devoted to collecting large data sets.
-Depending on the issue at hand, these data sets may contain highly sensitive information about individuals or organizations.
-It is thus of importance who gets to see these data and in what they are protected from third parties accessing them.
-There have been reported cases of researchers working with large corpora of texts containing peoples' complete medical record who sent these files indiscriminately around by email.
-A second aspect of privacy is how the data is collected in the first place.
-It is a question of ethical concern whether the individuals whose data is collected agree to their personal information being used.
-What the standards should be here isstill  far from clear as automated data-mining techniques gain popularity.
-
-A second point concerns the disruptive potential machine learning techniques have on labor markets.
-As ever more cognitive tasks can be automated, the economic landscape of service sector economies is bound undergo change.
-Some workers will without a doubt be disadvantaged by these changes.
-It must concern the inventors and implentors of such technology what happens here.
-
-A third difficult point is the so-called \emph{double use} problem.
-New technology's use is often underspecified and could be used for benefit of humanity or to its detriment.
-Chat-bots developed by researchers in automated natural language processing, to name a concrete example, can be used to assist people to access governmental health-care serices, yet they have reportedly been used by the Mexican government to disrupt political discussions deemed subversive on social media.
-\cite{leidner2017ethical}
-
-Finally, machine learning harbours the potential to amplify structural discrimination and systematically disadvantage people because of their race, gender or other social category.
-If the models which are trained algorithmically are fed biased data, then predictions or classifications they make will reflect that bias.
-Data processed by machine learning techniques often contains societal prejudice and bias, so that unquestioningly using them for decision making can severely reinforce society's existing biases.
-A related issue is that data available to the algorithms may systematically omit information relevant to what it is trying to uncover, as when minority groups are not adequately represented in the data.
-Yet another issue is that machine learning techniques may use useful patterns and relationships in the data that are really mirrorring historical factors of discrimination and marginalization.
-\cite{barocas2016big}
+With the very large data collected, stored and processed, issues of privacy arise naturally.
+Especially when data mining techniques are used to scrape the web, people's explicit consent to provide their data can easily go by the board.
+Another recurrent problem is the two-sided nature of any technology: it can be used for the good or for dubious purposes, as fraud or spying, for example.
+Finally, big data and ML approaches brings the risk of reenforcing structural bias along racial, gender or others social lines in society.
+Models trained on biased historical data, for example, have been used to determine prison sentences which lead to consistently biased outcomes against blacks and latinos in the United States of America, land of the brave and home of the free \cite{angwin2016machine}.
 
@@
 \paragraph{What is NLP?}
-Natural language processing is a field of research that uses computational tools to study natural language text or speech.
-The knowledge thus obtained can be used to develop computer systems to serve a wide variety of purposes.
+Natural language processing is a field of research that uses computational tools to build natural language text or speech applications.
+It is related to computational linguistic, an area of linguistic that uses computational techniques to get a scientific understanding of natural language.
+The knowledge thus obtained can be used in natural language processing to develop computer systems that serve a wide variety of purposes.
 Principal applications include translation between natural languages, artificial intelligence, user interfaces, like chat-bots, for example, summarization of text or speech recognition \cite{chowdhury2003natural}.
@@
 Distributional semantics is a linguistic theory of meaning looking to represent lexical meaning of expressions as a function of the context in which they appear.
-It theoretically foundation is the so-called \emph{distributional hypothesis} which holds that two expressions are similar in meaning, if they occur in similar contexts \cite{harris1954distributional}.
+Its theoretically foundation is the so-called \emph{distributional hypothesis} which holds that two expressions are similar in meaning, if they occur in similar contexts \cite{harris1954distributional}.
 %Althought it was originally developed as a distributional theory - in the sense that ...
@@
 
-The theory of distributional semantics has been used successfully in computational semantics, as it has become possible to obtain semantic representations immediately from natural language data .
+The theory of distributional semantics has been used successfully in computational semantics - an branch of computational lingiustics -, as it has become possible to obtain semantic representations immediately from natural language data.
 The theoretical framework of distributional semantics has become widely used and further developed as the advent and success of computational methods have enabled linguist to build large-scale models of linguistic meaning using digitalised corpora of text.
@@
 \paragraph{Relation: NLP - DS}
-Distributional semantics is concerned with a specific topic in the study of natural language, namely, meaning. It could thus be seen as one of many sub-fields of natural language processing research. 
+Distributional semantics is concerned with the specific topic of meaning in the study of natural language.
+If viewed from an application-oriented standpoint it could be seen as one of many sub-fields of natural language processing research.
+Indeed, as far as the present paper is concerned, it is precisely the application of distributional semantics in industry that warrants heightened attention.
 From a philosophical standpoint, it is notoriously difficult to come to grips with the concept of meaning .
@@
 -Why not ethics by design
+A first concern is the issue of privacy.
+Since very large amounts of data are required for these methods to be effective, great efforts are devoted to collecting large data sets which may contain highly sensitive information about individuals or organizations.
+It is thus of importance who gets to see these data and how they are protected from third parties accessing them.
+There have been reported cases of researchers working with large corpora of texts containing peoples' complete medical record who sent these files indiscriminately around by email.
+A second aspect of privacy is how the data is collected in the first place.
+It is a question of ethical concern whether the individuals whose data is collected agree to their personal information being used.
+What the standards should be here isstill  far from clear as automated data-mining techniques gain popularity.
+
+A second point concerns the disruptive potential machine learning techniques have on labor markets.
+As ever more cognitive tasks can be automated, the economic landscape of service sector economies is bound undergo change.
+Some workers will without a doubt be disadvantaged by these changes.
+It must concern the inventors and implentors of such technology what happens here.
+
+A third difficult point is the so-called \emph{double use} problem.
+New technology's use is often underspecified and could be used for benefit of humanity or to its detriment.
+Chat-bots developed by researchers in automated natural language processing, to name a concrete example, can be used to assist people to access governmental health-care serices, yet they have reportedly been used by the Mexican government to disrupt political discussions deemed subversive on social media.
+\cite{leidner2017ethical}
+
+Finally, machine learning harbours the potential to amplify structural discrimination and systematically disadvantage people because of their race, gender or other social category.
+If the models which are trained algorithmically are fed biased data, then predictions or classifications they make will reflect that bias.
+Data processed by machine learning techniques often contains societal prejudice and bias, so that unquestioningly using them for decision making can severely reinforce society's existing biases.
+A related issue is that data available to the algorithms may systematically omit information relevant to what it is trying to uncover, as when minority groups are not adequately represented in the data.
+Yet another issue is that machine learning techniques may use useful patterns and relationships in the data that are really mirrorring historical factors of discrimination and marginalization.
+\cite{barocas2016big}
+
 \section{Guidelines for Machine Learning}\hypertarget{sec2}{}
@@
 The authors of PAA urge creators to issue a Social Impact Statement wherein they detail the effect of their product has on society with respect to the five principles.
-In this way the principles are associated with best practices in the ...
-
+In this way the principles are associated with a set of best practices.
 \paragraph{5 Principles, quick recap of each of them}
-
-We will now briefly go over the five principles to explaint their content.
+We will now briefly go over the five principles - \emph{responsibility, explanability, accuracy, auditability and fairness} to explain their intention.
+Responsibility requires developers to take into account possible adverse effects of their algorithmic system. There should be designated roles within the organization.
 

### 2017-05-29 pds/Paper/paper_pds.tex
get shits saved again

@@
 \paragraph{Public Attention is in ML minor extent in NLP, not in DS}
-Many of the issues lately mentioned are of great importance to modern society at large.
-Consequently, with the rise of big data people working at the intersection of information technology and ethics have attempted to lay down foundations on how to handle these technologies.
-Many of these efforts have naturally been directed at the broad topic of machine learning in general.\cite{leidner2017ehtical} Only very recently the discussion has turned on machine learning used for analysis and modeling of language - often referred to as \emph{natural language processing}.
-Presumably because this is only one of many particular fields where machine learning is applied and the discussion of ethics natural language processing is bound pertain to less cases of applications throughout society, it took some time for the discussion to reach this field.
+Many of the issues lately mentioned are of great importance to modern society at large, especially as.
+Consequently, people working at the intersection of information technology and ethics have attempted to lay down foundations on how to handle these technologies.
+Many of these efforts have naturally been directed at the broad topic of machine learning in general.\cite{leidner2017ehtical} Only a few years ago the discussion has turned on the specific use of ML for analysis and modeling of language - often referred to as \emph{natural language processing}.
+Presumably because this is only one of many particular fields where machine learning is applied and the discussion of ethics natural language processing is bound to pertain to less cases of applications throughout society, it took some time for the discussion to reach this field.
 
-Recently, however, a discussion has opened up in a hitherto completely unmentioned domain.
+A recently as 2016, however, a discussion has opened up in a hitherto completely unmentioned domain.
 Machine learning techniques used for the computer-based study of language meaning - a field known as \emph{computational distributional semantics} - has come under scrutiny.
-Various studies have shown that these methods yield results which are in important ways biased against particular groups as, for example, women \cite{google} \cite{wagner2015s}.The realization that algorithms fed on big data come with a host of problematic issues has led to the emergence of a variety of principles, concrete guidelines and public as well as institutional scrutiny.
+Various studies have shown that these methods yield results which are in important ways biased against particular groups as, for example, women \cite{google} \cite{wagner2015s}.
+\textbf{Why this is important. Mention Google searches using semantic information.}
+
+The realization that algorithms fed on big data come with a host of problematic issues has led to the emergence of a variety of principles, concrete guidelines and public as well as institutional scrutiny.
 
@@
 In the present paper we want to draw on the existing efforts made by experts in the fields of ethics, computer science, public policy and law to extend these principles and guidelines to the field of distributional semantics.
-Specifically, we will use the \emph{Principles for Accountable Algorithms and a Social Impact Statement for Algorithms} (PAASISA) issued by the organization \emph{Fairness, Accountability and Transparency in Machine Learning} (FATML) - a consortium of researches and experts from private and public organizations working on machine learning.
-The PAASISA comprise five guiding principles: responsibility, explainability, accuracy, auditibility and fairness \cite{principles}.
+Specifically, we will use the \emph{Principles for Accountable Algorithms} (PAA) issued by the organization \emph{Fairness, Accountability and Transparency in Machine Learning} (FATML) - a consortium of researches and experts from private and public organizations working on machine learning.
+The PAA comprise five guiding principles: responsibility, explainability, accuracy, auditibility and fairness \cite{principles}.
 We will assess potential violation of techniques used in distributional semantics for each of these principles and, where applicable, issue guidelines to prevent their negative impact upon vulnerable groups.
@@
 Currently, we want to focus on aspects of discrimination and bias in the outcome of distributional semantics models for two reasons.
-For one, these are the least explored issues for distributional semantics.
-Many of the ideas, guidelines and principles concerning privacy, double use or economic impact are directly transferable from the general machine learning debate to any sub-field that similarly employs big data and algorithms, as does computational distributional semantics.
+For one, these are the least explored ethical issues for distributional semantics.
+Many of the ideas, guidelines and principles concerning privacy, double use or economic impact are directly transferable from the general machine learning debate to any sub-field that similarly employs big data and algorithms.
+Computational distributional semantics is no exception.
 Debates concerning privacy etc.
 can follow the same lines as the discussion to establish general purpose ethical principles for machine learning.
-The issue of discrimination and bias, however, has a special relation to distributional semantics because it occupies in a unique position of computational language research.
+The issue of discrimination and bias, however, has a special relation to distributional semantics because the field occupies a unique position of computational language research.
 The central concern of distributional semantics, as its name gives away, is the concept of meaning.
 As we will argue presently, this relates it in a special way to many sensitive topics in the general machine learning discussion.
-Put succinctly, it sits at the intersection of concerns pertaining to data selection, validity or questionability of computational methods employed, interpretation of results and usage thereof.
+Put succinctly, it sits at the intersection of concerns pertaining to data selection, validity or questionability of computational methods, interpretation of results and usage thereof.
 
@@
 Furthermore, many non-technical articles have been published outside of academic channels.
-be specific, mention FATML. use nlp ethic article to get overview.\\
+be specific, mention FATML. \\
+\textbf{use nlp ethic article to get overview. do this shits.}
 

### 2017-06-01 pds/Paper/paper_pds.tex
made progress. i hope

@@
 \documentclass{article}
-\title{Philosophy of Distributional Semantics\\
+\title{Philosophy of DS\\
 \large Best Practices for the Application of Computational Methods to Questions of Meaning}
@@
 The seriousness of the potential to disrespect people's privacy and integrity or systematically disadvantage specific groups has led to a recent debate concerning ethical guidelines and best practices for the general use of ML.
-Little has been said, however, about the proper use of algorithms in the specific field of Distributional Semantics (DS).
+Little has been said, however, about the proper use of algorithms in the specific field of DS (DS).
 Some researchers have called attention to possible dangers of using ML in DS, as it harbours the potential to reinforce harmful biases along dimensions of race and gender.
@@
 
-In the present essay, we will mostly be concerned with the last of these problematic issues surrounding machine learning.
+In the present essay, we will mostly be concerned with the last of these problematic issues surrounding ML.
 
@@
 Consequently, people working at the intersection of information technology and ethics have attempted to lay down foundations on how to handle these technologies.
-Many of these efforts have naturally been directed at the broad topic of machine learning in general.\cite{leidner2017ehtical} Only a few years ago the discussion has turned on the specific use of ML for analysis and modeling of language - often referred to as \emph{natural language processing}.
-Presumably because this is only one of many particular fields where machine learning is applied and the discussion of ethics natural language processing is bound to pertain to less cases of applications throughout society, it took some time for the discussion to reach this field.
+Many of these efforts have naturally been directed at the broad topic of ML in general.\cite{leidner2017ehtical} Only a few years ago the discussion has turned on the specific use of ML for analysis and modeling of language - often referred to as \emph{natural language processing}.
+Presumably because this is only one of many particular fields where ML is applied and the discussion of ethics  is bound to pertain to less cases of applications throughout society, it took some time for the discussion to reach this field.
 
 A recently as 2016, however, a discussion has opened up in a hitherto completely unmentioned domain.
-Machine learning techniques used for the computer-based study of language meaning - a field known as \emph{computational distributional semantics} - has come under scrutiny.
+ML techniques used for the computer-based study of language meaning - a field known as \emph{computational DS} - has come under scrutiny.
 Various studies have shown that these methods yield results which are in important ways biased against particular groups as, for example, women \cite{google} \cite{wagner2015s}.
@@
 \paragraph{Contribution: Drawing on ML discussion for DS}
-In the present paper we want to draw on the existing efforts made by experts in the fields of ethics, computer science, public policy and law to extend these principles and guidelines to the field of distributional semantics.
-Specifically, we will use the \emph{Principles for Accountable Algorithms} (PAA) issued by the organization \emph{Fairness, Accountability and Transparency in Machine Learning} (FATML) - a consortium of researches and experts from private and public organizations working on machine learning.
+In the present paper we want to draw on the existing efforts made by experts in the fields of ethics, computer science, public policy and law to extend these principles and guidelines to the field of DS.
+Specifically, we will use the \emph{Principles for Accountable Algorithms} (PAA) issued by the organization \emph{Fairness, Accountability and Transparency in Machine Learning} (FATML) - a consortium of researches and experts from private and public organizations working on ML.
 The PAA comprise five guiding principles: responsibility, explainability, accuracy, auditibility and fairness \cite{principles}.
-We will assess potential violation of techniques used in distributional semantics for each of these principles and, where applicable, issue guidelines to prevent their negative impact upon vulnerable groups.
+We will assess potential violation of techniques used in DS for each of these principles and, where applicable, issue guidelines to prevent their negative impact upon vulnerable groups.
 
-Currently, we want to focus on aspects of discrimination and bias in the outcome of distributional semantics models for two reasons.
-For one, these are the least explored ethical issues for distributional semantics.
-Many of the ideas, guidelines and principles concerning privacy, double use or economic impact are directly transferable from the general machine learning debate to any sub-field that similarly employs big data and algorithms.
-Computational distributional semantics is no exception.
+Currently, we want to focus on aspects of discrimination and bias in the outcome of DS models for two reasons.
+For one, these are the least explored ethical issues for DS.
+Many of the ideas, guidelines and principles concerning privacy, double use or economic impact are directly transferable from the general ML debate to any sub-field that similarly employs big data and algorithms.
+Computational DS is no exception.
 Debates concerning privacy etc.
-can follow the same lines as the discussion to establish general purpose ethical principles for machine learning.
-The issue of discrimination and bias, however, has a special relation to distributional semantics because the field occupies a unique position of computational language research.
-The central concern of distributional semantics, as its name gives away, is the concept of meaning.
-As we will argue presently, this relates it in a special way to many sensitive topics in the general machine learning discussion.
+can follow the same lines as the discussion to establish general purpose ethical principles for ML.
+The issue of discrimination and bias, however, has a special relation to DS because the field occupies a unique position of computational language research.
+The central concern of DS, as its name gives away, is the concept of meaning.
+As we will argue presently, this relates it in a special way to many sensitive topics in the general ML discussion.
 Put succinctly, it sits at the intersection of concerns pertaining to data selection, validity or questionability of computational methods, interpretation of results and usage thereof.
@@
 \paragraph{Methodoology}
-The method employed by the present contribution is one of scanning previous efforts in a similar vein, identifying useful parts and applying them to ethical questions surrounding distributional semantics.
+The method employed by the present contribution is one of scanning previous efforts in a similar vein, identifying useful parts and applying them to ethical questions surrounding DS.
 There will be no foundational efforts undertaken to sustain the validity of the principles thus transferred.
@@
 \paragraph{Paper Overview}
-The paper is structured as follows: In \hyperlink{sec2}{Section 2} we first give a very brief description of machine learning, natural language processing and distributional semantics as well as their relation to each other.
-Especially for distributional semantics we want to provide overview over societal domains where such technology is currently applied and could conceivably be applied in the future.
+The paper is structured as follows: In \hyperlink{sec2}{Section 2} we first give a very brief description of ML,  and DS as well as their relation to each other.
+Especially for DS we want to provide overview over societal domains where such technology is currently applied and could conceivably be applied in the future.
 \hyperlink{sec3}{Section 3} contains a discussion of the important ways in which DS, as a particular application of big data analysis, differs from other instances of ML.
@@
 \hyperlink{sec5}{Section 5} contains concluding remarks while \hyperlink{sec6}{Section 6} lists the references.
-\section{Machine Learning, Natural Language Processing and Distributional Semantics}\hypertarget{sec3}{ }
+\section{Machine Learning, Natural Language Processing and Distributional Semantics}\hypertarget{sec2}{ }
 \paragraph{What is ML?}
 The term \emph{Machine Learning} is an umbrella term for a variety of statistical methods that are able to \emph{learn} from a series of examples presented to them.
-The the set of examples - often called the \emph{training data} - is not simply memorized but the machine learning algorithm extracts generalizable knowledge \cite{domingos2012few}.
+The the set of examples - often called the \emph{training data} - is not simply memorized but the ML algorithm extracts generalizable knowledge \cite{domingos2012few}.
  What the algorithm learns are patterns and relationships within the training data and the result of the learning process is a statistical description - or \emph{model} - of the training data \cite{fayyad2001digital}.
@@
 It is related to computational linguistic, an area of linguistic that uses computational techniques to get a scientific understanding of natural language.
-The knowledge thus obtained can be used in natural language processing to develop computer systems that serve a wide variety of purposes.
+The knowledge thus obtained can be used in  to develop computer systems that serve a wide variety of purposes.
 Principal applications include translation between natural languages, artificial intelligence, user interfaces, like chat-bots, for example, summarization of text or speech recognition \cite{chowdhury2003natural}.
 \paragraph{Relation: ML - NLP}
-The area of natural language processing frequently serves itself of methods from machine learning, for example to process very large textual corpora.
-Indeed, parts of natural language processing can be thought of as particular applications of machine learning techniques to the domain of natural language.
-\paragraph{What is Distributional Semantics?}
-Distributional semantics is a linguistic theory of meaning looking to represent lexical meaning of expressions as a function of the context in which they appear.
+The area of  frequently serves itself of methods from ML, for example to process very large textual corpora.
+Indeed, parts of  can be thought of as particular applications of ML techniques to the domain of natural language.
+\paragraph{What is DS?}
+DS is a linguistic theory of meaning looking to represent lexical meaning of expressions as a function of the context in which they appear.
 Its theoretically foundation is the so-called \emph{distributional hypothesis} which holds that two expressions are similar in meaning, if they occur in similar contexts \cite{harris1954distributional}.
@@
 
-The theory of distributional semantics has been used successfully in computational semantics - an branch of computational lingiustics -, as it has become possible to obtain semantic representations immediately from natural language data.
-The theoretical framework of distributional semantics has become widely used and further developed as the advent and success of computational methods have enabled linguist to build large-scale models of linguistic meaning using digitalised corpora of text.
+The theory of DS has been used successfully in computational semantics - an branch of computational lingiustics -, as it has become possible to obtain semantic representations immediately from natural language data.
+The theoretical framework of DS has become widely used and further developed as the advent and success of computational methods have enabled linguist to build large-scale models of linguistic meaning using digitalised corpora of text.
 Following the distributional hypothesis, the meaning of an expression is a function of the context in which it occurs, where context can be construed in different ways.
@@
 \cite{boleda2016formal}
-Distributional semantics, then, takes the context in which an expression occurs and abstracts away from the particular contexts in a large corpus of natural language data.
+DS, then, takes the context in which an expression occurs and abstracts away from the particular contexts in a large corpus of natural language data.
 The mathematical vehicle for representing these abstractions are commonly vectors, but there are also other algebraic objects possible, as matrices or tensors.
@@
 \paragraph{Relation: NLP - DS}
-Distributional semantics is concerned with the specific topic of meaning in the study of natural language.
-If viewed from an application-oriented standpoint it could be seen as one of many sub-fields of natural language processing research.
-Indeed, as far as the present paper is concerned, it is precisely the application of distributional semantics in industry that warrants heightened attention.
+DS is concerned with the specific topic of meaning in the study of natural language.
+If viewed from an application-oriented standpoint it could be seen as one of many sub-fields of  research.
+Indeed, as far as the present paper is concerned, it is precisely the application of DS in industry that warrants heightened attention.
 From a philosophical standpoint, it is notoriously difficult to come to grips with the concept of meaning .
-Distributional semantics sits at the intersection of ethically touchy topics. 
+DS sits at the intersection of ethically touchy topics. 
 First off, consider the situation where a meaning space is constructed based on some seemingly innocuous corpus and ends up containing severe bias along lines of gender or racial stereotypes.
@@
 \paragraph{The state of the ethical discussion in ML, NLP and DS}
-For the domain general use of machine learning techniques there exists a series of studies by now from legal, political computer science experts which specify principles and guidelines.
+For the domain general use of ML techniques there exists a series of studies by now from legal, political computer science experts which specify principles and guidelines.
 Furthermore, many non-technical articles have been published outside of academic channels.
@@
 
-In the case of natural language processing there has been  has seen some movement in the last couple of years.
-For machine learning and natural language processing there are several  conferences on ethical issues by now.
+In the case of  there has been  has seen some movement in the last couple of years.
+For ML and  there are several  conferences on ethical issues by now.
 http://www.ethicsinnlp.org/related-conferences
@@
 
-A second point concerns the disruptive potential machine learning techniques have on labor markets.
+A second point concerns the disruptive potential ML techniques have on labor markets.
 As ever more cognitive tasks can be automated, the economic landscape of service sector economies is bound undergo change.
@@
 
-A third difficult point is the so-called \emph{double use} problem.
+A third difficult poit is the so-called \emph{double use} problem.
 New technology's use is often underspecified and could be used for benefit of humanity or to its detriment.
-Chat-bots developed by researchers in automated natural language processing, to name a concrete example, can be used to assist people to access governmental health-care serices, yet they have reportedly been used by the Mexican government to disrupt political discussions deemed subversive on social media.
+Chat-bots developed by researchers in automated , to name a concrete example, can be used to assist people to access governmental health-care serices, yet they have reportedly been used by the Mexican government to disrupt political discussions deemed subversive on social media.
 \cite{leidner2017ethical}
 
-Finally, machine learning harbours the potential to amplify structural discrimination and systematically disadvantage people because of their race, gender or other social category.
+Finally, ML harbours the potential to amplify structural discrimination and systematically disadvantage people because of their race, gender or other social category.
 If the models which are trained algorithmically are fed biased data, then predictions or classifications they make will reflect that bias.
-Data processed by machine learning techniques often contains societal prejudice and bias, so that unquestioningly using them for decision making can severely reinforce society's existing biases.
+Data processed by ML techniques often contains societal prejudice and bias, so that unquestioningly using them for decision making can severely reinforce society's existing biases.
 A related issue is that data available to the algorithms may systematically omit information relevant to what it is trying to uncover, as when minority groups are not adequately represented in the data.
-Yet another issue is that machine learning techniques may use useful patterns and relationships in the data that are really mirrorring historical factors of discrimination and marginalization.
+Yet another issue is that ML techniques may use useful patterns and relationships in the data that are really mirrorring historical factors of discrimination and marginalization.
 \cite{barocas2016big}
 
-\section{Guidelines for Machine Learning}\hypertarget{sec2}{}
+\section{Guidelines for Machine Learning}\hypertarget{sec3}{}
 \paragraph{Introduce FATML}
@@
 
-Given this state of affairs, it is important to hold the people who develop or implement such algorithmic systems accountable to the public. The FAT-ML concisely write in their PAA \cite{principles}
+Given this state of affairs, it is important to hold the people who develop or implement such algorithmic systems accountable to the public. The FATML concisely write in their PAA \cite{principles}
 \begin{quote}
@@
 \paragraph{5 Principles, quick recap of each of them}
-We will now briefly go over the five principles - \emph{responsibility, explanability, accuracy, auditability and fairness} to explain their intention.
-Responsibility requires developers to take into account possible adverse effects of their algorithmic system. There should be designated roles within the organization.
+We will now briefly go over the five principles - \emph{responsibility, explanability, accuracy, auditability and fairness} to describe their original intention in the PAA.
+Responsibility requires developers to take seriously the possibility that adverse effects are produced their algorithmic system.
+Consequently, there should be designated roles within the organization to be responsible for redress if harm is caused.
+Excuses akin to ''the algorithm did it'' are no valid scapegoats.
+This reflects the acknowledgment that, at some level, there are people who design these systems and put them to use.
+Explainability demands that both data selection and decisions taken by the algorithms can be explained to people affected by its usage in terms that are understandable by laywomen.
+The point here is not to obscure the workings of the systems to a point where no stakeholder can grasp it.
+Accuracy asks designers and users of the system to keep track of its mistakes and sources of error, so that worst case scenarios become more easily anticipited and possible damage reduced.
+Audibility is the requirement to create transparency in such a way that interested third parties are in the position to review the algorithms performance. This point is related to explainability.
+Finally, Fairness requires that the algorithmic system does not affect different demographic groups in disparate and unfavourable ways, that is, discriminates against them.
 
-\section{Application to DS}\hypertarget{sec4}{ }
-\paragraph{What's the same: --}
-\paragraph{What's different: Double-nature of textual corpora}
-\section{Conclusion}
+\paragraph{What's the same}
+It should be fair to say that some of these principles allow seamless transfer from general ML to any specific subdomain, as DS, without much further discussion.
+This goes for the principles corresponding to accuracy - loggin mistakes is good practice no matter where -, auditibility - one should never obstruct uninterested parties' inquiries into the possibility of things going wrong - and explainability - obfuscating one's work and keeping affected parties ignorant is evidently bad practice both in ML and in DS.
+
+\paragraph{What's different}
+Much more interesting than the overlap lately mentioned is the question how the two remaining principles apply to to DS.
+There are mainly two issues, one for each principle.
+
+Concerning the question of responsibility, consider a concrete scenario where the problem arises: Google trains a DS model on a corpus of news articles on Google News.
+The model contains semantic representations of lexical word meaning.
+Despite the articles in the corpus being written by professional journalists, these meanings contain clear bias along a binary gender dimension.
+One of many striking examples is the fact that the word \emph{woman} bears the same semantic relation to \emph{homemaker} in the meaning space as the word \emph{man} to \emph{computer scientist}.
+Now, a search engine like Google could use semantic information available in such a model, with bias and all, to optimize the relevance of their search results.
+It is quite conceivable that an employer queries Google for ''Computer Scientist Kabul'' to find professionals in the area and Google deems Linkedin profiles with male-associated names more relevant to the search request.
+Consequently, male computer scientists will find their profiles ranked higher in the results of the query and thus obtain more visibility.
+This would be a clear-cut case of discrimination and disadvantage based on gender, brought about directly by recommendation mechanism which utilizes DS.
+
+Who is to blame for this unfortunate outcome?
+The recommendation systems designers might simply point at the corpus in which they found their word meanings.
+After all, the corpus presumably reflects the fact that there are men than women who work in computer science accurately \cite{womenincs}.
+So the biased outcome of the search mirrors societal structure at, or so would the recommender's excuse sound.
+
+This little story goes to show that it can be excrutiatingly difficult to find the responsible human agent when DS is used for algorithmic systems.
+Even if one grants that the meanings as found in a corpus are biased in an unacceptable way, the question still remains what acceptability would entail.
+A follow to this is whether one can remove prejudice and bias from meaning spaces without rendering them unusable for usage in applications.
+What, for example, if the fact that the expression "woman" is functional in natural language discourse precisely because of its unfortunate semantic relationship to man.
+A recent paper by Bolukbasi et al.
+contradicts this pessimistic view \cite{bolukbasi2016man}.
+According to it, it is possible to debias the meanings of a semantic space along binary gender dimenions while still preserving its useful properties.
+
+The lesson of this particular successfull debiasing endeavour is straightforward.
+Performance is no general excuse for designers of systems employing DS to work with meanings that reflect societal bias.
+Efforts can be made to process the DS model after its original training to rid it of discriminatory disparaties.
+In general domain ML there is already a sizeable body of literature of finding an removing so-called \emph{disparate impact} - unintentional discrimination (cf.
+\cite{feldman2015certifying} \cite{dwork2012fairness}).
+
+Similar collections of standardized techniques should be developed to debias semantic spaces. Doing this is especially important, of course, if the DS model to be used for some outside application. 
+Perhaps a complement to the attempts to debiase embeddings, then, are the studies whose stated goal it is to uncover biases in corpora, as eg.
+\cite{wagner2015s} or \cite{herbelot2012distributional}.
+Findings that indicate clear instances of bias in widely used corpora can then be used as an indication where  debiasing would have to take place before the meanings should be put to further use.
+
+Both of these
+
+\section{Conclusion}\hypertarget{sec x}{}
 

### 2017-06-01 pds/Paper/pds_references.tex
get shits saved
made progress. i hope

@@
 
-@misc{principles,
+@misc{discrimination,
   title = {{Discrimination algorithms disparate impact} Kernel Description},
@@
   note = {Accessed: 2017-05-22}
+}
+@misc{womenincs,
+  title = {{Women in Computer Science} Kernel Description},
+  howpublished = {\url{http://www.computerscience.org/resources/women-in-computer-science/
+}},
+  note = {Accessed: 2017-06-1}
 }
@@
   pages =        {9--18},
+}
+
+@article{angwin2016machine,
+  title={Machine bias: There’s software used across the country to predict future criminals. and it’s biased against blacks},
+  author={Angwin, Julia and Larson, Jeff and Mattu, Surya and Kirchner, Lauren},
+  journal={ProPublica, May},
+  volume={23},
+  year={2016}
 }
@@
 }
+@inproceedings{herbelot2012distributional,
+  title={Distributional techniques for philosophical enquiry},
+  author={Herbelot, Aur{\'e}lie and Von Redecker, Eva and M{\"u}ller, Johanna},
+  booktitle={Proceedings of the 6th Workshop on Language Technology for Cultural Heritage, Social Sciences, and Humanities},
+  pages={45--54},
+  year={2012},
+  organization={Association for Computational Linguistics}
+}
 
@@
 }
+@inproceedings{dwork2012fairness,
+  title={Fairness through awareness},
+  author={Dwork, Cynthia and Hardt, Moritz and Pitassi, Toniann and Reingold, Omer and Zemel, Richard},
+  booktitle={Proceedings of the 3rd Innovations in Theoretical Computer Science Conference},
+  pages={214--226},
+  year={2012},
+  organization={ACM}
+}

### 2017-06-01 pds/Paper/paper_pds.tex
made progress. certainly. coke is the best.

@@
 The seriousness of the potential to disrespect people's privacy and integrity or systematically disadvantage specific groups has led to a recent debate concerning ethical guidelines and best practices for the general use of ML.
-Little has been said, however, about the proper use of algorithms in the specific field of DS (DS).
+Little has been said, however, about the proper use of algorithms in the specific field of Distributional Semantics (DS).
 Some researchers have called attention to possible dangers of using ML in DS, as it harbours the potential to reinforce harmful biases along dimensions of race and gender.
@@
 %1-2 paragraphs, talk about application, overall field
-\paragraph{Place of Machine Learning in society}
 As vast amounts of digital data have become available over past decades, algorithms are employed to use the data in order to take and inform decisions all across society. 
-Algorithms have a say in determining who is granted mortgages, who is released on bail or invited for job applications. 
-\cite{algorithms2016}
+Algorithms have a say in determining who is granted mortgages, who is released on bail or invited for job applications \cite{algorithms2016}.
 The theory behind many of these algorithms is \emph{machine learning} (ML).
@@
 
-\paragraph{Harmful Effects}
 There are several different sources of harmful effects that these techniques can bring about.
@@
 Finally, big data and ML approaches brings the risk of reenforcing structural bias along racial, gender or others social lines in society.
-Models trained on biased historical data, for example, have been used to determine prison sentences which lead to consistently biased outcomes against blacks and latinos in the United States of America, land of the brave and home of the free \cite{angwin2016machine}.
+Models trained on biased historical data, for example, have been used to determine prison sentences which lead to consistently biased outcomes against blacks and latinos in the United States of America \cite{angwin2016machine}.
 
-In the present essay, we will mostly be concerned with the last of these problematic issues surrounding ML.
+%largest part of intro, preview of paper, methodology!
+Clearly, many of these issues are of great importance to modern society at large.
+Consequently, people working at the intersection of information technology and ethics have attempted to lay down foundations on how to handle these technologies.
+The realization that algorithms fed on big data come with a host of problematic issues has led to the emergence of a variety of principles, concrete guidelines and public as well as institutional scrutiny.
+Many of these efforts have naturally been directed at the broad topic of ML in general.\cite{leidner2017ehtical} Only a few years ago the discussion has turned on the specific use of ML for analysis and modeling of language - sometimes broadly referred to as \emph{natural language processing} (NLP).
+This is only one of many particular fields where ML is applied and the discussion of ethics in ML naturally dissipated from more general to more domain-specific applications.
 
-%\begin{quote}
-%In the years since, several disparate impact cases have made their way to the Supreme Court and lower courts, most having to do with employment discrimination.
-This June, the Supreme Court’s decision in Texas Dept.
-of Housing and Community Affairs v.
-Inclusive Communities Project, Inc.
-affirmed the use of the disparate impact theory to fight housing discrimination.
-The Inclusive Communities Project had used a statistical analysis of housing patterns to show that a tax credit program effectively segregated Texans by race.
-Sorelle Friedler, a computer science
-%researcher at Haverford College and a fellow at Data \& Society, called the Court’s decision “huge,” both “in favor of civil rights…and in favor of statistics.”
-%\end{quote}
-
-%\paragraph{Contribution}
-%largest part of intro, preview of paper, methodology!
-\paragraph{Public Attention is in ML minor extent in NLP, not in DS}
-Many of the issues lately mentioned are of great importance to modern society at large, especially as.
-Consequently, people working at the intersection of information technology and ethics have attempted to lay down foundations on how to handle these technologies.
-Many of these efforts have naturally been directed at the broad topic of ML in general.\cite{leidner2017ehtical} Only a few years ago the discussion has turned on the specific use of ML for analysis and modeling of language - often referred to as \emph{natural language processing}.
-Presumably because this is only one of many particular fields where ML is applied and the discussion of ethics  is bound to pertain to less cases of applications throughout society, it took some time for the discussion to reach this field.
-
-A recently as 2016, however, a discussion has opened up in a hitherto completely unmentioned domain.
+A recently as 2016, however, a discussion has opened up in a hitherto completely undiscussed domain.
 ML techniques used for the computer-based study of language meaning - a field known as \emph{computational DS} - has come under scrutiny.
-Various studies have shown that these methods yield results which are in important ways biased against particular groups as, for example, women \cite{google} \cite{wagner2015s}.
+Various studies have shown that these methods can  yield results which are in important ways biased against particular groups as, for example, women \cite{google} \cite{wagner2015s}.
 \textbf{Why this is important. Mention Google searches using semantic information.}
 
-The realization that algorithms fed on big data come with a host of problematic issues has led to the emergence of a variety of principles, concrete guidelines and public as well as institutional scrutiny.
-
-\paragraph{Contribution: Drawing on ML discussion for DS}
-In the present paper we want to draw on the existing efforts made by experts in the fields of ethics, computer science, public policy and law to extend these principles and guidelines to the field of DS.
-Specifically, we will use the \emph{Principles for Accountable Algorithms} (PAA) issued by the organization \emph{Fairness, Accountability and Transparency in Machine Learning} (FATML) - a consortium of researches and experts from private and public organizations working on ML.
+In the present paper we want to draw on the existing efforts made by experts in the fields of ethics, computer science, public policy and law to extend principles  and guidelines for general purpose ML ethics to the field of DS.
+Specifically, we will use the \emph{Principles for Accountable Algorithms} (PAA) issued by the organization \emph{Fairness , Accountability and Transparency in Machine Learning} (FATML) - a consortium of researches and experts from private and public organizations working on ML.
 The PAA comprise five guiding principles: responsibility, explainability, accuracy, auditibility and fairness \cite{principles}.
 We will assess potential violation of techniques used in DS for each of these principles and, where applicable, issue guidelines to prevent their negative impact upon vulnerable groups. 
+The present contribution works by scanning previous efforts made in a similar vein, identifying useful parts therein and applying them to ethical questions surrounding DS.
+We will make no foundational efforts to sustain the validity of the principles.
+They are presupposed.
+Our analysis merely serves to discuss the usefulness of the principles in particular domain, and assert it where due.
 
-Currently, we want to focus on aspects of discrimination and bias in the outcome of DS models for two reasons.
-For one, these are the least explored ethical issues for DS.
+Currently, we want to focus on aspects of discrimination and bias in the outcome of DS models.
+This for two reasons: For one, these are the least explored ethical issues for DS.
 Many of the ideas, guidelines and principles concerning privacy, double use or economic impact are directly transferable from the general ML debate to any sub-field that similarly employs big data and algorithms.
@@
 
-\paragraph{Methodoology}
-The method employed by the present contribution is one of scanning previous efforts in a similar vein, identifying useful parts and applying them to ethical questions surrounding DS.
-There will be no foundational efforts undertaken to sustain the validity of the principles thus transferred.
-Their validity is presupposed.
-The analysis contained in this contribution merely serves to discuss the usefulness of the principles in particular domain, and assert it where due.
-
-\paragraph{Paper Overview}
 The paper is structured as follows: In \hyperlink{sec2}{Section 2} we first give a very brief description of ML,  and DS as well as their relation to each other.
@@
 For the present paper these particularities are important insofar as they entail differences when it comes to its risks and accountabilities across the spectrum of ML.
-In \hyperlink{sec4}{Section 4}, then, we try to apply the \emph{Principles for Accountable Algorithms and a Social Impact Statement for Algorithms} by FAT/ML to the specific case of DS.
+In \hyperlink{sec4}{Section 4}, then, we try to apply the \emph{Principles for Accountable Algorithms} by FATML to the specific case of DS.
 \hyperlink{sec5}{Section 5} contains concluding remarks while \hyperlink{sec6}{Section 6} lists the references.
-\section{Machine Learning, Natural Language Processing and Distributional Semantics}\hypertarget{sec2}{ }
-\paragraph{What is ML?}
-The term \emph{Machine Learning} is an umbrella term for a variety of statistical methods that are able to \emph{learn} from a series of examples presented to them.
-The the set of examples - often called the \emph{training data} - is not simply memorized but the ML algorithm extracts generalizable knowledge \cite{domingos2012few}.
- What the algorithm learns are patterns and relationships within the training data and the result of the learning process is a statistical description - or \emph{model} - of the training data \cite{fayyad2001digital}.
-\paragraph{What is NLP?}
-Natural language processing is a field of research that uses computational tools to build natural language text or speech applications.
+\section{Machine Learning and Distributional Semantics}\hypertarget{sec2}{ }
+The term \emph{machine learning} is an umbrella term for a variety of statistical methods that are able to \emph{learn} from a series of examples presented to them.
+The set of examples - often called the \emph{training data} - is not simply memorized, but the ML algorithm extracts generalizable knowledge like structures, patterns or relationships \cite{domingos2012few}.
+The result of the learning process is a statistical description - or \emph{model} - of the training data \cite{fayyad2001digital}.
+
+Natural language processing (NLP) is a field of research that uses computational tools to build natural language text or speech applications.
 It is related to computational linguistic, an area of linguistic that uses computational techniques to get a scientific understanding of natural language.
@@
 
+A further point to consider is the fact that the nature of the object supposedly captured by DS - meaning - is highly contested.
+Conceptual confusion and debate abound concerning the exact relation around the concepts of meaning, groundedness, speaker and community.
+This can make it even harder to adequately respond to calls of responsibility in these matters because the relation that holds between corpus, DS techniques, meaning and society are far from settled.
+
 This little story goes to show that it can be excrutiatingly difficult to find the responsible human agent when DS is used for algorithmic systems.
@@
 \cite{feldman2015certifying} \cite{dwork2012fairness}).
-
 Similar collections of standardized techniques should be developed to debias semantic spaces. Doing this is especially important, of course, if the DS model to be used for some outside application. 
@@
 \cite{wagner2015s} or \cite{herbelot2012distributional}.
-Findings that indicate clear instances of bias in widely used corpora can then be used as an indication where  debiasing would have to take place before the meanings should be put to further use.
+Findings that indicate clear instances of bias in widely used corpora can then be used as an indication where  debiasing should have to take place before the DS is put to further use.
+Thus, it appears that a two-fold effort is required in to prevent damage from happening in the employment of algorithmically generated meanings.
+On the one hand, one will constantly have to scan the semantic spaces gathered from corpora for undesired connotations or other semantic relationships within it; on the other, once such biases are certified, efforts need to be undertaken, so that systems using this semantic information do not systematically disadvantage people.
 
-Both of these
+In this fashion fairness principle could be addressed in DS.
+This does not yet respond, however, to the question where exactly responsibility lies - that is, with whom - in case something does go wrong.
 

### 2017-06-01 pds/Paper/paper_pds.tex
coca cola that is. shoutout to evo morales.

@@
 The present paper wants to point to one particular attempt to provide general-purpose guidelines for ML and draw on these existing insights to suggest guidelines and best practices for computational methods in DS.
-As the domain of DS differs in important aspects from other applications of ML, notably also for Natural Language Processing (NLP),
+As the domain of DS differs in important aspects from other applications of ML, notably also for natural language processing,
 this involves assessing  the applicability of general purpose ML guidelines to DS with these difference in mind.
@@
 Many of these efforts have naturally been directed at the broad topic of ML in general.\cite{leidner2017ehtical} Only a few years ago the discussion has turned on the specific use of ML for analysis and modeling of language - sometimes broadly referred to as \emph{natural language processing} (NLP).
-This is only one of many particular fields where ML is applied and the discussion of ethics in ML naturally dissipated from more general to more domain-specific applications.
+This is only one of many particular fields where ML is applied and the discussion of ethics in ML naturally seep through from more general to more domain-specific applications.
 
@@
 
-The paper is structured as follows: In \hyperlink{sec2}{Section 2} we first give a very brief description of ML,  and DS as well as their relation to each other.
+The paper is structured as follows: In \hyperlink{sec2}{Section 2} we first give a very brief description of ML, NLP and DS as well as their relation to each other.
 Especially for DS we want to provide overview over societal domains where such technology is currently applied and could conceivably be applied in the future.
@@
 \hyperlink{sec5}{Section 5} contains concluding remarks while \hyperlink{sec6}{Section 6} lists the references.
-\section{Machine Learning and Distributional Semantics}\hypertarget{sec2}{ }
+\section{Machine Learning, Natural Language Processing and Distributional Semantics}\hypertarget{sec2}{ }
+The three terms that make up the title of this section are frequently muddled in literature pertaining to what this paper is about.
+In order to avoid any confusion, and give people foreign to any such field a quick overview, the following paragraphs contain a short description of the three terms and their mutual relation.
+
 The term \emph{machine learning} is an umbrella term for a variety of statistical methods that are able to \emph{learn} from a series of examples presented to them.
@@
 
-Natural language processing (NLP) is a field of research that uses computational tools to build natural language text or speech applications.
+\emph{Natural language processing} is a field of research that uses computational tools to build natural language text or speech applications.
 It is related to computational linguistic, an area of linguistic that uses computational techniques to get a scientific understanding of natural language.
-The knowledge thus obtained can be used in  to develop computer systems that serve a wide variety of purposes.
+The knowledge thus obtained can be used  to develop computer systems that serve a wide variety of purposes.
 Principal applications include translation between natural languages, artificial intelligence, user interfaces, like chat-bots, for example, summarization of text or speech recognition \cite{chowdhury2003natural}.
-\paragraph{Relation: ML - NLP}
-The area of  frequently serves itself of methods from ML, for example to process very large textual corpora.
-Indeed, parts of  can be thought of as particular applications of ML techniques to the domain of natural language.
-\paragraph{What is DS?}
-DS is a linguistic theory of meaning looking to represent lexical meaning of expressions as a function of the context in which they appear.
+The area of NLP frequently serves itself of methods from ML, for example to process very large textual corpora.
+Indeed, parts of NLP can be thought of as particular applications of ML techniques to the domain of natural language.
+
+\emph{Distributional semantics} is a linguistic theory of meaning looking to represent lexical meaning of expressions as a function of the context in which they appear.
 Its theoretically foundation is the so-called \emph{distributional hypothesis} which holds that two expressions are similar in meaning, if they occur in similar contexts \cite{harris1954distributional}.
-%Althought it was originally developed as a distributional theory - in the sense that ...
-- in the middle of the 20th century, it can be seen as following the footsteps of structural linguists, notably Bloomfield \cite{bloomfield}
 
-The theory of DS has been used successfully in computational semantics - an branch of computational lingiustics -, as it has become possible to obtain semantic representations immediately from natural language data.
+The theory of DS has been used successfully in computational semantics - a branch of computational lingiustics.
+This recent success is attributable to increased availability of large amounts of digitalised language data and ML techniqeus.Particularly, it has become possible to obtain semantic representations immediately from natural language data.
 The theoretical framework of DS has become widely used and further developed as the advent and success of computational methods have enabled linguist to build large-scale models of linguistic meaning using digitalised corpora of text.
+
 Following the distributional hypothesis, the meaning of an expression is a function of the context in which it occurs, where context can be construed in different ways.
@@
 
-\paragraph{Relation: NLP - DS}
 DS is concerned with the specific topic of meaning in the study of natural language.

### 2017-06-01 pds/Paper/paper_pds.tex
final commit of day

@@
 The theory of DS has been used successfully in computational semantics - a branch of computational lingiustics.
-This recent success is attributable to increased availability of large amounts of digitalised language data and ML techniqeus.Particularly, it has become possible to obtain semantic representations immediately from natural language data.
+This recent success is attributable to increased availability of large amounts of digitalised language data and ML techniqeus.
+Particularly, it has become possible to obtain semantic representations immediately from natural language data.
 The theoretical framework of DS has become widely used and further developed as the advent and success of computational methods have enabled linguist to build large-scale models of linguistic meaning using digitalised corpora of text.
@@
 DS is concerned with the specific topic of meaning in the study of natural language.
-If viewed from an application-oriented standpoint it could be seen as one of many sub-fields of  research.
+If viewed from an application-oriented standpoint it could be seen as one of many sub-fields of NLP research.
 Indeed, as far as the present paper is concerned, it is precisely the application of DS in industry that warrants heightened attention.
-From a philosophical standpoint, it is notoriously difficult to come to grips with the concept of meaning .
-DS sits at the intersection of ethically touchy topics. 
-First off, consider the situation where a meaning space is constructed based on some seemingly innocuous corpus and ends up containing severe bias along lines of gender or racial stereotypes.
-A recent study has found, for example, that a model trained on Google News articles display significant bias along binary gender dimensions \cite{bolukbasi2016man}.
+For conreteness' sake, consider a situation where a meaning space is constructed based on some seemingly innocuous corpus and ends up containing severe bias along lines of gender or racial stereotypes.
+A recent study has found, for example, that a model trained on Google News articles displays significant bias along binary gender dimensions \cite{bolukbasi2016man}.
 This raises a host of questions concerning the provenance of the bias, the adequate response by the responsible distributional semanticist and, not to mention, the nature of meaning.
@@
 After all, the corpus was produced by a particular society, so it should be expected that it mirror that society's prejudices.
-Is the stance "the model picked up what it is supposed to pick up, blame society" a fair one, or has the computational linguist the moral obligation to rid her model of biased meaning representations?
+Is the stance "the model picked up what it is supposed to pick up, blame society" a fair one, or has whoever works with the model the obligation to rid it of biased meaning representations?
+The questions go on: If the semantic model is further used for some interaction within that same society that taught it - via corpus - what it means to be a \emph{woman}, will the bias be further re-enforced?
+It is has been reported that when one queries Google with the name of a person, then Google is more likely to show advertisement related to that person being arrested if the name's connotation is black \cite{sweeney2013discrimination}.
 
-Fair and square, the model contains bias because the corpus contains bias.
-What if the semantic model is further used for some interaction within that same society that taught it - via corpus - what it means to be a \emph{woman}, will the bias be further re-enforced?
-It is has been reported that when one queries Google with the name of a person, then Google is more likely to show advertisement related to that person being arrested if the name's connotation is black \cite{sweeney2013discrimination}.
-%A related issue is that even if results are valid, the of over-, or, respectively, under-representation of specific items in the data can disadvantage group.
-
+\section{Ethics of Machine Learning}
 \paragraph{The state of the ethical discussion in ML, NLP and DS}
-For the domain general use of ML techniques there exists a series of studies by now from legal, political computer science experts which specify principles and guidelines.
+For the domain general use of ML techniques there exists a series of studies by from legal, political computer science experts which specify principles and guidelines.
 Furthermore, many non-technical articles have been published outside of academic channels.
@@
 
-\section{Guidelines for Machine Learning}\hypertarget{sec3}{}
-\paragraph{Introduce FATML}
+\section{Principles for  for Machine Learning}\hypertarget{sec3}{}
 Algorithmically informed decisions are used in ever more domains of the private and public sectors and thus have the potential to impact society in significant ways.
@@
 In this way the principles are associated with a set of best practices.
-\paragraph{5 Principles, quick recap of each of them}
+
 We will now briefly go over the five principles - \emph{responsibility, explanability, accuracy, auditability and fairness} to describe their original intention in the PAA.
@@
 
-\paragraph{What's the same}
 It should be fair to say that some of these principles allow seamless transfer from general ML to any specific subdomain, as DS, without much further discussion.
 This goes for the principles corresponding to accuracy - loggin mistakes is good practice no matter where -, auditibility - one should never obstruct uninterested parties' inquiries into the possibility of things going wrong - and explainability - obfuscating one's work and keeping affected parties ignorant is evidently bad practice both in ML and in DS.
-
-\paragraph{What's different}
-Much more interesting than the overlap lately mentioned is the question how the two remaining principles apply to to DS.
-There are mainly two issues, one for each principle.
+Much more interesting than the overlap lately mentioned, however, is the question how the two remaining principles apply to to DS.
 

### 2017-06-02 pds/Paper/paper_pds.tex
cleanup intro

@@
 There are several different sources of harmful effects that these techniques can bring about.
-With the very large data collected, stored and processed, issues of privacy arise naturally.
-Especially when data mining techniques are used to scrape the web, people's explicit consent to provide their data can easily go by the board.
+With the very large amounts of data collected, stored and processed, issues of privacy arise naturally.
+Especially when sophisticated data gathering techniques are used to scrape the web, people's explicit consent to provide their data can easily go by the board.
 Another recurrent problem is the two-sided nature of any technology: it can be used for the good or for dubious purposes, as fraud or spying, for example.
-Finally, big data and ML approaches brings the risk of reenforcing structural bias along racial, gender or others social lines in society.
-Models trained on biased historical data, for example, have been used to determine prison sentences which lead to consistently biased outcomes against blacks and latinos in the United States of America \cite{angwin2016machine}.
+Finally, big data and ML approaches harbour the risk of re-enforcing structures of discrimination along racial, gender or others social lines.
+In a widely publicized case, to name one concrete example, automated systems have been used for legal decisions which lead to consistently biased outcomes against blacks and latinos in the United States of America \cite{angwin2016machine}.
 
-%largest part of intro, preview of paper, methodology!
-Clearly, many of these issues are of great importance to modern society at large.
+Many of these issues are of great importance to modern society at large.
 Consequently, people working at the intersection of information technology and ethics have attempted to lay down foundations on how to handle these technologies.
-The realization that algorithms fed on big data come with a host of problematic issues has led to the emergence of a variety of principles, concrete guidelines and public as well as institutional scrutiny.
-Many of these efforts have naturally been directed at the broad topic of ML in general.\cite{leidner2017ehtical} Only a few years ago the discussion has turned on the specific use of ML for analysis and modeling of language - sometimes broadly referred to as \emph{natural language processing} (NLP).
-This is only one of many particular fields where ML is applied and the discussion of ethics in ML naturally seep through from more general to more domain-specific applications.
+The realization that algorithms fed on big data come with a host of problematic issues has led increased public as well as institutional scrutiny (notably, the oval office \cite{united2014big}) and, consequently, to the elaboration of a variety of principles and concrete guidelines for their use.
+Many of these efforts have naturally been directed at the broad topic of ML in general \cite{leidner2017ehtical}.
+Only a few years ago the discussion has turned on the specific use of ML for analysis and modeling of language - sometimes broadly referred to as \emph{natural language processing} (NLP).
 
-A recently as 2016, however, a discussion has opened up in a hitherto completely undiscussed domain.
-ML techniques used for the computer-based study of language meaning - a field known as \emph{computational DS} - has come under scrutiny.
+As recently as 2016, however, a discussion has opened up in a hitherto completely undiscussed domain.
+ML techniques used for the computer-based study of language meaning - a field known as \emph{computational distributional semantics} - has come under scrutiny.
 Various studies have shown that these methods can  yield results which are in important ways biased against particular groups as, for example, women \cite{google} \cite{wagner2015s}.
-\textbf{Why this is important. Mention Google searches using semantic information.}
+This form of machine bias in word meaning can affect people in discriminatory ways.
+It is quite conceivable that in the near future many applications, for example automated customer services, will operate with automatically generated representations of natural language meaning.
+If semantic bias translates to bias into discriminatory decisions being taken, then there is a problem.
 
-In the present paper we want to draw on the existing efforts made by experts in the fields of ethics, computer science, public policy and law to extend principles  and guidelines for general purpose ML ethics to the field of DS.
+In the present paper we want to draw on the existing efforts made by experts in the fields of ethics, computer science, public policy and law to extend principles  and guidelines for general purpose ML ethics to the field of distributional semantics (DS).
 Specifically, we will use the \emph{Principles for Accountable Algorithms} (PAA) issued by the organization \emph{Fairness , Accountability and Transparency in Machine Learning} (FATML) - a consortium of researches and experts from private and public organizations working on ML.
@@
 We will assess potential violation of techniques used in DS for each of these principles and, where applicable, issue guidelines to prevent their negative impact upon vulnerable groups.
+
 The present contribution works by scanning previous efforts made in a similar vein, identifying useful parts therein and applying them to ethical questions surrounding DS.
@@
 Our analysis merely serves to discuss the usefulness of the principles in particular domain, and assert it where due.
-
-Currently, we want to focus on aspects of discrimination and bias in the outcome of DS models.
+Currently, we want to focus on aspects of discrimination and bias in the outcome of DS.
 This for two reasons: For one, these are the least explored ethical issues for DS.
@@
 As we will argue presently, this relates it in a special way to many sensitive topics in the general ML discussion.
-Put succinctly, it sits at the intersection of concerns pertaining to data selection, validity or questionability of computational methods, interpretation of results and usage thereof.
+Put succinctly, it sits at the intersection of concerns pertaining to data selection, validity or questionability of computational methods, interpretation of its results and the use it is given in applications.
 

### 2017-06-02 pds/Paper/paper_pds.tex
nearly done. gotta cut.

@@
 \usepackage{amsmath}
+\usepackage{biblatex}
+\addbibresource{pds_references.bib}
+\usepackage{csquotes}
 
-%\usepackage{biblatex}
-%\addbibresource{pds_references.bib}
-\usepackage{csquotes}
 \begin{document}
@@
 Little has been said, however, about the proper use of algorithms in the specific field of Distributional Semantics (DS).
-Some researchers have called attention to possible dangers of using ML in DS, as it harbours the potential to reinforce harmful biases along dimensions of race and gender.
+Some researchers have called attention to possible dangers of using ML in DS, as it harbors the potential to reinforce harmful biases along dimensions of race and gender.
 The present paper wants to point to one particular attempt to provide general-purpose guidelines for ML and draw on these existing insights to suggest guidelines and best practices for computational methods in DS.
@@
 \section{Introduction}
-%1-2 paragraphs, talk about application, overall field
 As vast amounts of digital data have become available over past decades, algorithms are employed to use the data in order to take and inform decisions all across society. 
@@
 Another recurrent problem is the two-sided nature of any technology: it can be used for the good or for dubious purposes, as fraud or spying, for example.
-Finally, big data and ML approaches harbour the risk of re-enforcing structures of discrimination along racial, gender or others social lines.
-In a widely publicized case, to name one concrete example, automated systems have been used for legal decisions which lead to consistently biased outcomes against blacks and latinos in the United States of America \cite{angwin2016machine}.
+Finally, big data and ML approaches harbor the risk of re-enforcing structures of discrimination along racial, gender or others social lines.
+In a widely publicized case, to name one concrete example, automated systems have been used for legal decisions which lead to consistently biased outcomes against Blacks and Latinos in the United States of America \cite{angwin2016machine}.
 
@@
 The realization that algorithms fed on big data come with a host of problematic issues has led increased public as well as institutional scrutiny (notably, the oval office \cite{united2014big}) and, consequently, to the elaboration of a variety of principles and concrete guidelines for their use.
-Many of these efforts have naturally been directed at the broad topic of ML in general \cite{leidner2017ehtical}.
+Many of these efforts have naturally been directed at the broad topic of ML in general \cite{leidner2017ethical}.
 Only a few years ago the discussion has turned on the specific use of ML for analysis and modeling of language - sometimes broadly referred to as \emph{natural language processing} (NLP).
@@
 ML techniques used for the computer-based study of language meaning - a field known as \emph{computational distributional semantics} - has come under scrutiny.
-Various studies have shown that these methods can  yield results which are in important ways biased against particular groups as, for example, women \cite{google} \cite{wagner2015s}.
+Various studies have shown that these methods can  yield results which are in important ways biased against particular groups as, for example, women \cite{bolukbasi2016man} \cite{wagner2015s}.
 This form of machine bias in word meaning can affect people in discriminatory ways.
@@
 Specifically, we will use the \emph{Principles for Accountable Algorithms} (PAA) issued by the organization \emph{Fairness , Accountability and Transparency in Machine Learning} (FATML) - a consortium of researches and experts from private and public organizations working on ML.
-The PAA comprise five guiding principles: responsibility, explainability, accuracy, auditibility and fairness \cite{principles}.
-We will assess potential violation of techniques used in DS for each of these principles and, where applicable, issue guidelines to prevent their negative impact upon vulnerable groups.
+The PAA comprise five guiding principles: responsibility, explainability, accuracy, auditability and fairness \cite{principles}.
+We will assess whether techniques used in DS could potentially violate these principles. This then allows us to issue guidelines to prevent their negative impact upon vulnerable groups.
 
@@
 As we will argue presently, this relates it in a special way to many sensitive topics in the general ML discussion.
-Put succinctly, it sits at the intersection of concerns pertaining to data selection, validity or questionability of computational methods, interpretation of its results and the use it is given in applications.
+Put succinctly, it sits at the intersection of concerns pertaining to data selection, validity or questionableness of computational methods, interpretation of its results and the use it is given in applications.
 
 The paper is structured as follows: In \hyperlink{sec2}{Section 2} we first give a very brief description of ML, NLP and DS as well as their relation to each other.
-Especially for DS we want to provide overview over societal domains where such technology is currently applied and could conceivably be applied in the future.
-\hyperlink{sec3}{Section 3} contains a discussion of the important ways in which DS, as a particular application of big data analysis, differs from other instances of ML.
-For the present paper these particularities are important insofar as they entail differences when it comes to its risks and accountabilities across the spectrum of ML.
-In \hyperlink{sec4}{Section 4}, then, we try to apply the \emph{Principles for Accountable Algorithms} by FATML to the specific case of DS.
-\hyperlink{sec5}{Section 5} contains concluding remarks while \hyperlink{sec6}{Section 6} lists the references.
+Also contained in this section is an assessment of the extent in which ehtical discussions are already led in this field.
+\hyperlink{sec3}{Section 3} sketches in what ways ML is susceptible to harm people and introduces the PAA.
+This forms the backdrop for \hyperlink{sec4}{Section 4} where we assess in what precise way this specific set of ethical principles designed for ML is applicable to DS.
+\hyperlink{sec5}{Section 5} contains concluding remarks and \hyperlink{sec6}{Section 6} lists the references.
+
 \section{Machine Learning, Natural Language Processing and Distributional Semantics}\hypertarget{sec2}{ }
-The three terms that make up the title of this section are frequently muddled in literature pertaining to what this paper is about.
-In order to avoid any confusion, and give people foreign to any such field a quick overview, the following paragraphs contain a short description of the three terms and their mutual relation.
+The three terms that make up the title of this section are frequently muddled in related literature.
+In order to avoid confusion, and give people alien to any such field a quick overview, the following paragraphs contain a short description of the three terms and their relation.
 
@@
 The theory of DS has been used successfully in computational semantics - a branch of computational lingiustics.
-This recent success is attributable to increased availability of large amounts of digitalised language data and ML techniqeus.
+This recent success is attributable to increased availability of large amounts of digitalized language data and ML techniqeus.
 Particularly, it has become possible to obtain semantic representations immediately from natural language data.
@@
 Context is not restricted to such simple analysis, however, as it can possibly contain more complex syntactic relations within a text, or even visual, textual or auditory information gathered from data.
-\cite{boleda2016formal}
-DS, then, takes the context in which an expression occurs and abstracts away from the particular contexts in a large corpus of natural language data.
+
+DS, then, takes the contexts in which an expression occurs in a large corpus of natural language data and abstracts away from the particular contexts.
 The mathematical vehicle for representing these abstractions are commonly vectors, but there are also other algebraic objects possible, as matrices or tensors.
 The meaning of an expression, as a function of its context, is then represented by the distribution across the dimensions of the vector.
-A set of words thus represented builds a semantic space - mathematically, a vector space - wherein one can study the semantic relationship between the expressions because some semantic relationships between expressions are mirrored by geometric relationship in these spaces.
+A set of words thus represented builds a semantic space - mathematically, a vector space - wherein one can study the semantic relationship between the expressions because some semantic relationships between expressions are mirrored by geometric relationship in these spaces \cite{boleda2016formal}.
 
@@
 Indeed, as far as the present paper is concerned, it is precisely the application of DS in industry that warrants heightened attention.
-For conreteness' sake, consider a situation where a meaning space is constructed based on some seemingly innocuous corpus and ends up containing severe bias along lines of gender or racial stereotypes.
+Consider a situation where a meaning space is constructed based on some seemingly innocuous corpus and ends up containing severe bias along lines of gender or racial stereotypes.
 A recent study has found, for example, that a model trained on Google News articles displays significant bias along binary gender dimensions \cite{bolukbasi2016man}.
@@
 
-Does one have to live with the fact that the meanings discovered in some corpus systematically contain unfavourable connotations for some group of people?
+Does one have to live with the fact that the meanings discovered in some corpus systematically contain unfavorable connotations for some group of people?
 After all, the corpus was produced by a particular society, so it should be expected that it mirror that society's prejudices.
-Is the stance "the model picked up what it is supposed to pick up, blame society" a fair one, or has whoever works with the model the obligation to rid it of biased meaning representations?
-The questions go on: If the semantic model is further used for some interaction within that same society that taught it - via corpus - what it means to be a \emph{woman}, will the bias be further re-enforced?
-It is has been reported that when one queries Google with the name of a person, then Google is more likely to show advertisement related to that person being arrested if the name's connotation is black \cite{sweeney2013discrimination}.
+What position should one assume towards these results?
+Is it fair to claim that the the model picked up what it is supposed to pick up and blame society for systematically biased meanings?
+Or does whoever works with the model the obligation to rid it of biased meaning representations?
+Such questions gain concrete importance if the semantic model is further used for some interaction within that same society that taught it - via corpus - what it means to be a \emph{woman}, for example.
+What if using the model could lead to re-enforcing the structures that lead to the biased corpus in the first place?
+These are not mere hypotheticals, but already very concrete issues.
+It is has been reported that when one queries Google with the name of a person, then Google is more likely to show advertisement related to that person's criminal record if the name is associated with black people \cite{sweeney2013discrimination}.
 
-\section{Ethics of Machine Learning}
-\paragraph{The state of the ethical discussion in ML, NLP and DS}
-For the domain general use of ML techniques there exists a series of studies by from legal, political computer science experts which specify principles and guidelines.
+The great import of these issues seems to be recognized by some people in the field of ML.
+At least of the domain-general use of ML techniques there exists a series of studies by from legal, political computer science experts which specify principles and guidelines (cf.
+\cite{pasquale2015black}\cite{feldman2015certifying}\cite{barocas2016big}.
+Starting in 2014 with a workshop series at the \emph{Conference on Neural Information Processing Systems}, an organization called \emph{Fairness, Accountability, and Transparency in Machine Learning} (FATML) has been working on solutions for ethical problems in ML.
 Furthermore, many non-technical articles have been published outside of academic channels.
-be specific, mention FATML. \\
-\textbf{use nlp ethic article to get overview. do this shits.}
 
-In the case of  there has been  has seen some movement in the last couple of years.
-For ML and  there are several  conferences on ethical issues by now.
-http://www.ethicsinnlp.org/related-conferences
+In the case of NLP there has been  has seen some movement in the last couple of years.
+There have been discussion regarding the moral agency of chat-bots \cite{thieltges2016devil}, as well as more general contributions to ethical guidelines in broader NLP \cite{hovy2016social}.
+Additionally, both for ML and NLP many public events were hosted and conferences established to discuss ethical issues (cf. \cite{conferences}).
+As a field, however, DS in particular has seen little attention in this respect.
+Although there are already proposals for best practices in NLP \cite{leidner2017ethical}, for example, no similar effort has been made for DS.
+\section{Principles for  for Machine Learning}\hypertarget{sec3}{}
+Algorithmically informed decisions are used in ever more domains of the private and public sectors and thus have the potential to impact society in significant ways.
+We now specify four ways in which unreflected usage of ML can bring about harm.
 
- even concrete proposals for best practices.\\
--DS in particular has seen little attention.
-
-\paragraph{What good should come from looking at DS?}
--focus not on dual-use or bad treatment of mechanical turks, but the question how bias in meaning is dangerous.\\
--Why not ethics by design
 A first concern is the issue of privacy.
@@
 It is a question of ethical concern whether the individuals whose data is collected agree to their personal information being used.
-What the standards should be here isstill  far from clear as automated data-mining techniques gain popularity.
 
 A second point concerns the disruptive potential ML techniques have on labor markets.
-As ever more cognitive tasks can be automated, the economic landscape of service sector economies is bound undergo change.
+As ever more cognitive tasks can be automated, the economic landscape of service sector economies is bound to undergo big change.
 Some workers will without a doubt be disadvantaged by these changes.
-It must concern the inventors and implentors of such technology what happens here.
 
-A third difficult poit is the so-called \emph{double use} problem.
+A third difficult point is the so-called \emph{double use} problem.
 New technology's use is often underspecified and could be used for benefit of humanity or to its detriment.
-Chat-bots developed by researchers in automated , to name a concrete example, can be used to assist people to access governmental health-care serices, yet they have reportedly been used by the Mexican government to disrupt political discussions deemed subversive on social media.
-\cite{leidner2017ethical}
+Chat-bots developed by researchers in NLP, to name a concrete example, can be used to assist people to access governmental health-care services, yet they have reportedly been used by the Mexican government to disrupt political discussions deemed subversive on social media \cite{leidner2017ethical}.
 
-Finally, ML harbours the potential to amplify structural discrimination and systematically disadvantage people because of their race, gender or other social category.
+Finally, ML harbors the potential to amplify structural discrimination and systematically disadvantage people because of their race, gender or other social category.
 If the models which are trained algorithmically are fed biased data, then predictions or classifications they make will reflect that bias.
-Data processed by ML techniques often contains societal prejudice and bias, so that unquestioningly using them for decision making can severely reinforce society's existing biases.
+Data processed by ML techniques can reflect societal prejudice and bias, so that unquestioningly using them for decision making can reinforce society's existing discriminatory structure.
 A related issue is that data available to the algorithms may systematically omit information relevant to what it is trying to uncover, as when minority groups are not adequately represented in the data.
-Yet another issue is that ML techniques may use useful patterns and relationships in the data that are really mirrorring historical factors of discrimination and marginalization.
-\cite{barocas2016big}
+Yet another issue is that ML techniques may use useful patterns and relationships in the data that are really mirroring historical factors of discrimination and marginalization \cite{barocas2016big}.
 
-\section{Principles for  for Machine Learning}\hypertarget{sec3}{}
-Algorithmically informed decisions are used in ever more domains of the private and public sectors and thus have the potential to impact society in significant ways.
 In many cases of undesirable effects that come with these technologies, legislation is behind in adequately protecting people negatively affected by it.
-
-Given this state of affairs, it is important to hold the people who develop or implement such algorithmic systems accountable to the public. The FATML concisely write in their PAA \cite{principles}
+Given this state of affairs, it is important to hold the people who develop or implement such algorithmic systems accountable to the public. The FATML concisely write in their PAA: 
 \begin{quote}
@@
 ''The algorithm did it'' is not an acceptable excuse if algorithmic systems make mistakes or have undesired consequences, including from machine-learning processes.
-\end{quote}
+\end{quote}\cite{principles}
 
-The document then goes on to name five principles that should guide developders of algorithms in designing and implementing their systems.
+The document then goes on to name five principles that should guide developers of algorithms in designing and implementing their systems.
 The authors of PAA urge creators to issue a Social Impact Statement wherein they detail the effect of their product has on society with respect to the five principles.
@@
 
-We will now briefly go over the five principles - \emph{responsibility, explanability, accuracy, auditability and fairness} to describe their original intention in the PAA.
+We will now briefly go over the five principles - \emph{responsibility, explainability, accuracy, auditability and fairness} to describe their original intention in the PAA.
 Responsibility requires developers to take seriously the possibility that adverse effects are produced their algorithmic system.
@@
 The point here is not to obscure the workings of the systems to a point where no stakeholder can grasp it.
-Accuracy asks designers and users of the system to keep track of its mistakes and sources of error, so that worst case scenarios become more easily anticipited and possible damage reduced.
+Accuracy asks designers and users of the system to keep track of its mistakes and sources of error, so that worst case scenarios become more easily anticipated and possible damage reduced.
 Audibility is the requirement to create transparency in such a way that interested third parties are in the position to review the algorithms performance. This point is related to explainability.
-Finally, Fairness requires that the algorithmic system does not affect different demographic groups in disparate and unfavourable ways, that is, discriminates against them.
-
-It should be fair to say that some of these principles allow seamless transfer from general ML to any specific subdomain, as DS, without much further discussion.
-This goes for the principles corresponding to accuracy - loggin mistakes is good practice no matter where -, auditibility - one should never obstruct uninterested parties' inquiries into the possibility of things going wrong - and explainability - obfuscating one's work and keeping affected parties ignorant is evidently bad practice both in ML and in DS.
+Finally, Fairness requires that the algorithmic system does not affect different demographic groups in disparate and unfavorable ways, that is, discriminates against them.
+We now want to assess to what degree we can use these principles for the field of DS.
+\section{Principles for Distributional Semantics}\hypertarget{sec x}{}
+It should be fair to say that some of these principles allow seamless transfer from general ML to any specific sub-domain, as DS, without much further discussion.
+This goes for the principles corresponding to accuracy - loging mistakes is good practice no matter where -, auditability - one should never obstruct uninterested parties' inquiries into the possibility of things going wrong - and explainability - obfuscating one's work and keeping affected parties ignorant is evidently bad practice both in ML and in DS.
 Much more interesting than the overlap lately mentioned, however, is the question how the two remaining principles apply to to DS.
@@
 Now, a search engine like Google could use semantic information available in such a model, with bias and all, to optimize the relevance of their search results.
-It is quite conceivable that an employer queries Google for ''Computer Scientist Kabul'' to find professionals in the area and Google deems Linkedin profiles with male-associated names more relevant to the search request.
+It is quite conceivable that an employer queries Google for ''Computer Scientist Kabul'' to find professionals in the area and Google deems LinkedIn profiles with male-associated names more relevant to the search request.
 Consequently, male computer scientists will find their profiles ranked higher in the results of the query and thus obtain more visibility.
@@
 
-This little story goes to show that it can be excrutiatingly difficult to find the responsible human agent when DS is used for algorithmic systems.
+This little story goes to show that it can be excruciatingly difficult to find the responsible human agent when DS is used for algorithmic systems.
 Even if one grants that the meanings as found in a corpus are biased in an unacceptable way, the question still remains what acceptability would entail.
@@
 contradicts this pessimistic view \cite{bolukbasi2016man}.
-According to it, it is possible to debias the meanings of a semantic space along binary gender dimenions while still preserving its useful properties.
+According to it, it is possible to debias the meanings of a semantic space along binary gender dimensions while still preserving its useful properties.
 
-The lesson of this particular successfull debiasing endeavour is straightforward.
+The lesson of this particular successfull debiasing endeavor is straightforward.
 Performance is no general excuse for designers of systems employing DS to work with meanings that reflect societal bias.
 Efforts can be made to process the DS model after its original training to rid it of discriminatory disparaties.
-In general domain ML there is already a sizeable body of literature of finding an removing so-called \emph{disparate impact} - unintentional discrimination (cf.
+In general domain ML there is already a sizable body of literature of finding an removing so-called \emph{disparate impact} - unintentional discrimination (cf.
 \cite{feldman2015certifying} \cite{dwork2012fairness}).
@@
 This does not yet respond, however, to the question where exactly responsibility lies - that is, with whom - in case something does go wrong.
-
+It is perhaps the most intricate difficulty concerning ethics in DS as potential culprits are society at large, the producers of the corpus, the distributional semanticist or the person who implements the system that interacts with society again.
+The nature of language meaning and how language and society interact are notoriously contested - the heated debates about what is appropriate in gendering language serves as a case in point.
+For data engineers and programmers who work with these systems there is certainly no easy answer to this issue at first sight.
+Perhaps the best that can be offered in terms of practical advice is that if the possibilty exists to algorithmically rid a DS model of biases potentially relevant to its application, then this should be done.
 \section{Conclusion}\hypertarget{sec x}{}
@@
 \section{References}\hypertarget{sec5}{ }
-%\printbibliography
+\printbibliography
 

### 2017-06-02 pds/Paper/paper_pds.tex
errthang, but concl.

@@
 \documentclass{article}
-\title{Philosophy of DS\\
-\large Best Practices for the Application of Computational Methods to Questions of Meaning}
+\title{Practical Ethics for Computational Semantics\\
+\large Philosophy of Distributional Semantics}
 \date{}
@@
 \usepackage{csquotes}
-
 \begin{document}
@@
 The present paper wants to point to one particular attempt to provide general-purpose guidelines for ML and draw on these existing insights to suggest guidelines and best practices for computational methods in DS.
-As the domain of DS differs in important aspects from other applications of ML, notably also for natural language processing,
+As the domain of DS differs in important aspects from other applications of ML, notably also natural language processing,
 this involves assessing  the applicability of general purpose ML guidelines to DS with these difference in mind.
@@
 With the very large amounts of data collected, stored and processed, issues of privacy arise naturally.
-Especially when sophisticated data gathering techniques are used to scrape the web, people's explicit consent to provide their data can easily go by the board.
+Especially when sophisticated data gathering techniques are used to scrape the web, peoples's explicit consent to provide their data can easily go by the board.
 Another recurrent problem is the two-sided nature of any technology: it can be used for the good or for dubious purposes, as fraud or spying, for example.
 Finally, big data and ML approaches harbor the risk of re-enforcing structures of discrimination along racial, gender or others social lines.
-In a widely publicized case, to name one concrete example, automated systems have been used for legal decisions which lead to consistently biased outcomes against Blacks and Latinos in the United States of America \cite{angwin2016machine}.
+In a widely publicized case, to name one concrete example, automated systems have been used for legal decisions which lead to consistently biased outcomes against Blacks and Latinos in the USA \cite{angwin2016machine}.
 
-Many of these issues are of great importance to modern society at large.
+Many of these issues are of great importance to modern societies.
 Consequently, people working at the intersection of information technology and ethics have attempted to lay down foundations on how to handle these technologies.
@@
 ML techniques used for the computer-based study of language meaning - a field known as \emph{computational distributional semantics} - has come under scrutiny.
-Various studies have shown that these methods can  yield results which are in important ways biased against particular groups as, for example, women \cite{bolukbasi2016man} \cite{wagner2015s}.
+Various studies have demonstrated that these methods can  yield results which are in important ways biased against particular groups as, for example, women \cite{bolukbasi2016man}\cite{wagner2015s}.
 This form of machine bias in word meaning can affect people in discriminatory ways.
 It is quite conceivable that in the near future many applications, for example automated customer services, will operate with automatically generated representations of natural language meaning.
-If semantic bias translates to bias into discriminatory decisions being taken, then there is a problem.
+If semantic bias translates into discriminatory decisions being taken, then there is a problem.
 
@@
 The PAA comprise five guiding principles: responsibility, explainability, accuracy, auditability and fairness \cite{principles}.
-We will assess whether techniques used in DS could potentially violate these principles. This then allows us to issue guidelines to prevent their negative impact upon vulnerable groups.
+We will assess in what ways techniques used in DS could potentially violate these principles. This then allows us to issue guidelines to prevent their negative impact upon vulnerable groups.
 
@@
 They are presupposed.
-Our analysis merely serves to discuss the usefulness of the principles in particular domain, and assert it where due.
+Our analysis merely serves to discuss the usefulness of the principles in particular domain, and assert them where due.
 Currently, we want to focus on aspects of discrimination and bias in the outcome of DS.
-This for two reasons: For one, these are the least explored ethical issues for DS.
-Many of the ideas, guidelines and principles concerning privacy, double use or economic impact are directly transferable from the general ML debate to any sub-field that similarly employs big data and algorithms.
+This is for two reasons: For one, these are the least explored ethical issues for DS.
+Furthermore, some of the ideas, guidelines and principles concerning privacy, abuse of otherwise legitimate technology ot  or economic impact are directly transferable from the general ML debate to any sub-field that similarly employs big data and algorithms.
 Computational DS is no exception.
@@
 
-\section{Machine Learning, Natural Language Processing and Distributional Semantics}\hypertarget{sec2}{ }
-The three terms that make up the title of this section are frequently muddled in related literature.
+\section{ML, NLP and DS}\hypertarget{sec2}{ }
+The three abbreviations that make up the title of this section are frequently muddled in related literature.
 In order to avoid confusion, and give people alien to any such field a quick overview, the following paragraphs contain a short description of the three terms and their relation.
@@
 The theory of DS has been used successfully in computational semantics - a branch of computational lingiustics.
-This recent success is attributable to increased availability of large amounts of digitalized language data and ML techniqeus.
+This recent success is attributable to increased availability of large amounts of digitalized language data and ML techniques.
 Particularly, it has become possible to obtain semantic representations immediately from natural language data.
@@
 Context is not restricted to such simple analysis, however, as it can possibly contain more complex syntactic relations within a text, or even visual, textual or auditory information gathered from data.
+\textbf{Moar refs?}
 
@@
 If viewed from an application-oriented standpoint it could be seen as one of many sub-fields of NLP research.
-Indeed, as far as the present paper is concerned, it is precisely the application of DS in industry that warrants heightened attention.
-Consider a situation where a meaning space is constructed based on some seemingly innocuous corpus and ends up containing severe bias along lines of gender or racial stereotypes.
-A recent study has found, for example, that a model trained on Google News articles displays significant bias along binary gender dimensions \cite{bolukbasi2016man}.
-This raises a host of questions concerning the provenance of the bias, the adequate response by the responsible distributional semanticist and, not to mention, the nature of meaning.
+Indeed, as far as the present paper is concerned, it is precisely the application of DS in industry that requires heightened attention.
 
-Does one have to live with the fact that the meanings discovered in some corpus systematically contain unfavorable connotations for some group of people?
-After all, the corpus was produced by a particular society, so it should be expected that it mirror that society's prejudices.
-What position should one assume towards these results?
-Is it fair to claim that the the model picked up what it is supposed to pick up and blame society for systematically biased meanings?
-Or does whoever works with the model the obligation to rid it of biased meaning representations?
-Such questions gain concrete importance if the semantic model is further used for some interaction within that same society that taught it - via corpus - what it means to be a \emph{woman}, for example.
-What if using the model could lead to re-enforcing the structures that lead to the biased corpus in the first place?
-These are not mere hypotheticals, but already very concrete issues.
-It is has been reported that when one queries Google with the name of a person, then Google is more likely to show advertisement related to that person's criminal record if the name is associated with black people \cite{sweeney2013discrimination}.
-
-The great import of these issues seems to be recognized by some people in the field of ML.
-At least of the domain-general use of ML techniques there exists a series of studies by from legal, political computer science experts which specify principles and guidelines (cf.
-\cite{pasquale2015black}\cite{feldman2015certifying}\cite{barocas2016big}.
-Starting in 2014 with a workshop series at the \emph{Conference on Neural Information Processing Systems}, an organization called \emph{Fairness, Accountability, and Transparency in Machine Learning} (FATML) has been working on solutions for ethical problems in ML.
-Furthermore, many non-technical articles have been published outside of academic channels.
-
-In the case of NLP there has been  has seen some movement in the last couple of years.
-There have been discussion regarding the moral agency of chat-bots \cite{thieltges2016devil}, as well as more general contributions to ethical guidelines in broader NLP \cite{hovy2016social}.
-Additionally, both for ML and NLP many public events were hosted and conferences established to discuss ethical issues (cf. \cite{conferences}).
-As a field, however, DS in particular has seen little attention in this respect.
-Although there are already proposals for best practices in NLP \cite{leidner2017ethical}, for example, no similar effort has been made for DS.
 \section{Principles for  for Machine Learning}\hypertarget{sec3}{}
@@
 
+In the broader field of ML some people seem to have recognized the importance ethical considerations.
+For the domain-general use of ML techniques there exists a series of studies by from legal, political computer science experts which specify principles and guidelines (cf.
+\cite{pasquale2015black}\cite{feldman2015certifying}\cite{barocas2016big}.
+Starting in 2014 with a workshop series at the \emph{Conference on Neural Information Processing Systems}, an organization called \emph{Fairness, Accountability, and Transparency in Machine Learning} (FATML) has been working on solutions for ethical problems in ML.
+Furthermore, many non-technical articles have been published outside of academic channels.
+
+In the case of NLP there has been  has seen some movement in the last couple of years.
+There have been discussion regarding the moral agency of chat-bots \cite{thieltges2016devil}, as well as more general contributions to ethical guidelines in broader NLP \cite{hovy2016social}.
+Additionally, both for ML and NLP many public events were hosted and conferences established to discuss ethical issues (cf. \cite{conferences}).
+As a field, however, DS in particular has seen little attention in this respect.
+Although there are already proposals for best practices in NLP \cite{leidner2017ethical}, for example, no similar effort has been made for DS.
+
 In many cases of undesirable effects that come with these technologies, legislation is behind in adequately protecting people negatively affected by it.
-Given this state of affairs, it is important to hold the people who develop or implement such algorithmic systems accountable to the public. The FATML concisely write in their PAA: 
+Given this state of affairs, it is important to hold the people who develop or implement such algorithmic systems accountable to the public. The FATML concisely write in their PAA\cite{principles}: 
 \begin{quote}
@@
 ''The algorithm did it'' is not an acceptable excuse if algorithmic systems make mistakes or have undesired consequences, including from machine-learning processes.
-\end{quote}\cite{principles}
+\end{quote}
 
@@
 We now want to assess to what degree we can use these principles for the field of DS.
-\section{Principles for Distributional Semantics}\hypertarget{sec x}{}
+\section{Principles for Distributional Semantics}\hypertarget{sec4}{}
 It should be fair to say that some of these principles allow seamless transfer from general ML to any specific sub-domain, as DS, without much further discussion.
@@
 
-Concerning the question of responsibility, consider a concrete scenario where the problem arises: Google trains a DS model on a corpus of news articles on Google News.
-The model contains semantic representations of lexical word meaning.
-Despite the articles in the corpus being written by professional journalists, these meanings contain clear bias along a binary gender dimension.
-One of many striking examples is the fact that the word \emph{woman} bears the same semantic relation to \emph{homemaker} in the meaning space as the word \emph{man} to \emph{computer scientist}.
-Now, a search engine like Google could use semantic information available in such a model, with bias and all, to optimize the relevance of their search results.
-It is quite conceivable that an employer queries Google for ''Computer Scientist Kabul'' to find professionals in the area and Google deems LinkedIn profiles with male-associated names more relevant to the search request.
+Concerning the question of fairness, consider a concrete scenario where the problem arises: A recent study has found, for example, that a model of lexical word meaning trained on Google News articles displays significant bias along binary gender dimensions \cite{bolukbasi2016man}.
+Despite the articles in the corpus being written by professional journalists, the word \emph{woman} bears the same semantic relation to \emph{homemaker} in the meaning space as the word \emph{man} to \emph{computer scientist}.
+A search engine could use semantic information available in such a model, with bias and all, to optimize the relevance of their search results.
+Imagine an employer querying Google for ''Computer Scientist Kabul'' to find professionals in the area and Google deems LinkedIn profiles with male-associated names more relevant to the search request.
 Consequently, male computer scientists will find their profiles ranked higher in the results of the query and thus obtain more visibility.
 This would be a clear-cut case of discrimination and disadvantage based on gender, brought about directly by recommendation mechanism which utilizes DS.
+These are not mere hypotheticals though, but already very concrete issues.
+It is has been reported that when one queries Google with the name of a person, then Google is more likely to show advertisement related to that person's criminal record if the name is associated with black people \cite{sweeney2013discrimination}.
 
+This raises a host of questions concerning the provenance of the bias in the model, the adequate response by the responsible distributional semanticist and, not to mention, the nature of meaning.
+Does one have to live with the fact that the meanings discovered in some corpus systematically contain unfavorable connotations for some group of people?
+After all, the corpus was produced by a particular society, so it should be expected that it mirror that society's prejudices.
+What position should one assume towards these results?
+Is it fair to claim that the the model picked up what it is supposed to pick up and blame society for systematically biased meanings?
+Or does whoever works with the model the obligation to rid it of biased meaning representations?
+Such questions gain concrete importance if the semantic model is further used for some interaction within that same society that taught it - via corpus - what it means to be a \emph{woman}, for example.
+What if using the model results in re-enforcement of the structures that lead to the biased corpus in the first place? 
 Who is to blame for this unfortunate outcome?
 The recommendation systems designers might simply point at the corpus in which they found their word meanings.
-After all, the corpus presumably reflects the fact that there are men than women who work in computer science accurately \cite{womenincs}.
-So the biased outcome of the search mirrors societal structure at, or so would the recommender's excuse sound.
+After all, the corpus accurately reflects the fact that there are more men than women who work in computer science \cite{womenincs}.
+So the biased outcome of the search mirrors societal structure, or so would the recommender's excuse might sound.
 
@@
 
-This little story goes to show that it can be excruciatingly difficult to find the responsible human agent when DS is used for algorithmic systems.
+This story goes to show that it can be excruciatingly difficult to find the responsible human agent when DS is used for algorithmic systems.
 Even if one grants that the meanings as found in a corpus are biased in an unacceptable way, the question still remains what acceptability would entail.
 A follow to this is whether one can remove prejudice and bias from meaning spaces without rendering them unusable for usage in applications.
-What, for example, if the fact that the expression "woman" is functional in natural language discourse precisely because of its unfortunate semantic relationship to man.
+What, for example, if the fact that the expression "woman" is functional in natural language discourse precisely because of its unfortunate semantic relationship to "man".
 A recent paper by Bolukbasi et al.
@@
 Performance is no general excuse for designers of systems employing DS to work with meanings that reflect societal bias.
-Efforts can be made to process the DS model after its original training to rid it of discriminatory disparaties.
+Efforts can be made to process the DS model after its original training to rid it of discriminatory disparities.
 In general domain ML there is already a sizable body of literature of finding an removing so-called \emph{disparate impact} - unintentional discrimination (cf.
@@
 Similar collections of standardized techniques should be developed to debias semantic spaces. Doing this is especially important, of course, if the DS model to be used for some outside application. 
-Perhaps a complement to the attempts to debiase embeddings, then, are the studies whose stated goal it is to uncover biases in corpora, as eg.
-\cite{wagner2015s} or \cite{herbelot2012distributional}.
+Perhaps a complement to the attempts to debiase embeddings, then, are the studies whose stated goal it is to uncover biases in corpora (cf.
+\cite{wagner2015s} \cite{herbelot2012distributional}).
 Findings that indicate clear instances of bias in widely used corpora can then be used as an indication where  debiasing should have to take place before the DS is put to further use.
@@
 For data engineers and programmers who work with these systems there is certainly no easy answer to this issue at first sight.
-Perhaps the best that can be offered in terms of practical advice is that if the possibilty exists to algorithmically rid a DS model of biases potentially relevant to its application, then this should be done.
-\section{Conclusion}\hypertarget{sec x}{}
+The best that can be offered in terms of practical advice is that if the possibilty exists to algorithmically rid a DS model of biases relevant to its application, then this should be done.
+\section{Conclusion}\hypertarget{sec5}{}
 
-\section{References}\hypertarget{sec5}{ }
+\section{References}\hypertarget{sec6}{ }
 \printbibliography

### 2017-06-02 pds/Paper/paper_pds.tex
last commit otd

@@
 It should be fair to say that some of these principles allow seamless transfer from general ML to any specific sub-domain, as DS, without much further discussion.
-This goes for the principles corresponding to accuracy - loging mistakes is good practice no matter where -, auditability - one should never obstruct uninterested parties' inquiries into the possibility of things going wrong - and explainability - obfuscating one's work and keeping affected parties ignorant is evidently bad practice both in ML and in DS.
+This goes for the principles corresponding to accuracy - logging mistakes is good practice no matter where -, auditability - one should never obstruct uninterested parties' inquiries into the possibility of things going wrong - and explainability - obfuscating one's work and keeping affected parties ignorant is evidently bad practice both in ML and in DS.
 Much more interesting than the overlap lately mentioned, however, is the question how the two remaining principles apply to to DS.
@@
 \section{Conclusion}\hypertarget{sec5}{}
-
+The preceding discussion seems to warrant but two clear take-aways for DS.
+First, if a semantic space does contain bias and will be further used in some application, then efforts should be undertaken to debias it.
+Second, if no the DS model cannot be debiased, then it should not be employed externally without further efforts to ensure that it does not discriminate.
+The responsibility for these two points ultimately lie with the designer of the application.
 \section{References}\hypertarget{sec6}{ }

### 2017-06-05 pds/Paper/paper_pds.tex
approaching the end

@@
 Some researchers have called attention to possible dangers of using ML in DS, as it harbors the potential to reinforce harmful biases along dimensions of race and gender.
-The present paper wants to point to one particular attempt to provide general-purpose guidelines for ML and draw on these existing insights to suggest guidelines and best practices for computational methods in DS.
-As the domain of DS differs in important aspects from other applications of ML, notably also natural language processing,
-this involves assessing  the applicability of general purpose ML guidelines to DS with these difference in mind.
+The present paper wants draw on one particular attempt to provide general-purpose guidelines for ML - the \emph{Principles for Accountable Algorithms} to suggest guidelines and best practices for computational methods in DS.
+As the phenomena of interest in DS differs in important aspects from other applications of ML, notably also natural language processing,
+this involves assessing the applicability of general purpose ML guidelines to DS.
+Our main finding is that most principles carry over smoothly, except for issues concerning responsibility and algorithmic fairness.
 \end{abstract}
@@
 There are several different sources of harmful effects that these techniques can bring about.
-With the very large amounts of data collected, stored and processed, issues of privacy arise naturally.
-Especially when sophisticated data gathering techniques are used to scrape the web, peoples's explicit consent to provide their data can easily go by the board.
+%With the very large amounts of data collected, stored and processed, issues of privacy arise naturally.
+Especially when sophisticated data gathering techniques are used to scrape the web, peoples' explicit consent to provide their data can easily go by the board.
 Another recurrent problem is the two-sided nature of any technology: it can be used for the good or for dubious purposes, as fraud or spying, for example.
 Finally, big data and ML approaches harbor the risk of re-enforcing structures of discrimination along racial, gender or others social lines.
-In a widely publicized case, to name one concrete example, automated systems have been used for legal decisions which lead to consistently biased outcomes against Blacks and Latinos in the USA \cite{angwin2016machine}.
-
-Many of these issues are of great importance to modern societies.
+%In a widely publicized case, to name one concrete example, automated systems have been used for legal decisions which lead to consistently biased outcomes against Blacks and Latinos in the USA \cite{angwin2016machine}.
+Problems include privacy concerns, abuse of technology or unintended structural discrimination; all are of great importance to modern societies.
 Consequently, people working at the intersection of information technology and ethics have attempted to lay down foundations on how to handle these technologies.
@@
 The area of NLP frequently serves itself of methods from ML, for example to process very large textual corpora.
-Indeed, parts of NLP can be thought of as particular applications of ML techniques to the domain of natural language.
+%Indeed, parts of NLP can be thought of as particular applications of ML techniques to the domain of natural language.
 
@@
 Most typically, it is other words surrounding an expression in a text - the sentence in which it appears, for example.
-Context is not restricted to such simple analysis, however, as it can possibly contain more complex syntactic relations within a text, or even visual, textual or auditory information gathered from data.
-\textbf{Moar refs?}
+Context is not restricted to such simple analysis, however, as it can possibly contain more complex syntactic relations within a text, or even visual, textual or auditory information gathered from data \cite{erk2012vector}.
 
@@
 
-DS is concerned with the specific topic of meaning in the study of natural language.
+%DS is concerned with the specific topic of meaning in the study of natural language.
 If viewed from an application-oriented standpoint it could be seen as one of many sub-fields of NLP research.
-Indeed, as far as the present paper is concerned, it is precisely the application of DS in industry that requires heightened attention.
-
-\section{Principles for  for Machine Learning}\hypertarget{sec3}{}
+% Indeed, as far as the present paper is concerned, it is precisely the application of DS in industry that requires heightened attention.
+\section{Principles for Machine Learning}\hypertarget{sec3}{}
 Algorithmically informed decisions are used in ever more domains of the private and public sectors and thus have the potential to impact society in significant ways.
@@
 A first concern is the issue of privacy.
-Since very large amounts of data are required for these methods to be effective, great efforts are devoted to collecting large data sets which may contain highly sensitive information about individuals or organizations.
+Since effective use of these methods requires very large amounts of data, great efforts are devoted to collecting data sets which may contain highly sensitive information about individuals or organizations.
 It is thus of importance who gets to see these data and how they are protected from third parties accessing them.
-There have been reported cases of researchers working with large corpora of texts containing peoples' complete medical record who sent these files indiscriminately around by email.
-A second aspect of privacy is how the data is collected in the first place.
-It is a question of ethical concern whether the individuals whose data is collected agree to their personal information being used.
+%There have been reported cases of researchers working with large corpora of texts containing peoples' complete medical record who sent these files indiscriminately around by email. 
+Another aspect of privacy is how the data is collected in the first place. Principally, whether the individuals whose data is collected agree to their personal information being used.
 
 A second point concerns the disruptive potential ML techniques have on labor markets.
-As ever more cognitive tasks can be automated, the economic landscape of service sector economies is bound to undergo big change.
+As ever more cognitive tasks can be automated, the economic landscape of service sector economies is bound to undergo big change \cite{west2015happens}.
 Some workers will without a doubt be disadvantaged by these changes.
 
-A third difficult point is the so-called \emph{double use} problem.
+A third difficulty is the so-called \emph{double use} problem.
 New technology's use is often underspecified and could be used for benefit of humanity or to its detriment.
@@
 A related issue is that data available to the algorithms may systematically omit information relevant to what it is trying to uncover, as when minority groups are not adequately represented in the data.
-Yet another issue is that ML techniques may use useful patterns and relationships in the data that are really mirroring historical factors of discrimination and marginalization \cite{barocas2016big}.
+Yet another issue is that ML techniques may exploit useful patterns and relationships in the data that are really mirroring historical factors of discrimination and marginalization \cite{barocas2016big}.
 
-In the broader field of ML some people seem to have recognized the importance ethical considerations.
-For the domain-general use of ML techniques there exists a series of studies by from legal, political computer science experts which specify principles and guidelines (cf.
+In the broader field of ML some people seem to have recognized the importance of ethical considerations.
+For the domain-general use of ML techniques there exists a series of studies by legal, political and ethical scholars, as well as computer science experts which specify principles and guidelines (cf.
 \cite{pasquale2015black}\cite{feldman2015certifying}\cite{barocas2016big}.
-Starting in 2014 with a workshop series at the \emph{Conference on Neural Information Processing Systems}, an organization called \emph{Fairness, Accountability, and Transparency in Machine Learning} (FATML) has been working on solutions for ethical problems in ML.
-Furthermore, many non-technical articles have been published outside of academic channels.
+Starting in 2014 with a workshop series at the \emph{Conference on Neural Information Processing Systems}, the organization FATML has been working on solutions for ethical problems in ML.
+Furthermore, many non-technical articles have been published outside of academic channels (cf.
+\cite{new2016}\cite{algorithms2016}).
 
-In the case of NLP there has been  has seen some movement in the last couple of years.
-There have been discussion regarding the moral agency of chat-bots \cite{thieltges2016devil}, as well as more general contributions to ethical guidelines in broader NLP \cite{hovy2016social}.
+In the case of NLP there has been some movement in the last couple of years with discussions regarding the moral agency of chat-bots \cite{thieltges2016devil}, as well as more general contributions to ethical guidelines in broader NLP \cite{hovy2016social}.
 Additionally, both for ML and NLP many public events were hosted and conferences established to discuss ethical issues (cf. \cite{conferences}).
@@
 
-We will now briefly go over the five principles - \emph{responsibility, explainability, accuracy, auditability and fairness} to describe their original intention in the PAA.
+We will now briefly go over the five principles - \emph{responsibility, explainability, accuracy, auditability} and \emph{fairness} to describe their original intention in the PAA.
 Responsibility requires developers to take seriously the possibility that adverse effects are produced their algorithmic system.
@@
 Imagine an employer querying Google for ''Computer Scientist Kabul'' to find professionals in the area and Google deems LinkedIn profiles with male-associated names more relevant to the search request.
-Consequently, male computer scientists will find their profiles ranked higher in the results of the query and thus obtain more visibility.
+Consequently, male computer scientists' profiles are ranked higher in the results and thus more visible.
 This would be a clear-cut case of discrimination and disadvantage based on gender, brought about directly by recommendation mechanism which utilizes DS.
@@
 Even if one grants that the meanings as found in a corpus are biased in an unacceptable way, the question still remains what acceptability would entail.
-A follow to this is whether one can remove prejudice and bias from meaning spaces without rendering them unusable for usage in applications.
-What, for example, if the fact that the expression "woman" is functional in natural language discourse precisely because of its unfortunate semantic relationship to "man".
+A follow-up question to this is whether one can remove prejudice and bias from meaning spaces without rendering them unusable for usage in applications.
+What, for example, if the fact that the expression \emph{woman} is functional in natural language discourse precisely because of its unfortunate semantic relationship to \emph{man}.
 A recent paper by Bolukbasi et al.
@@
 For data engineers and programmers who work with these systems there is certainly no easy answer to this issue at first sight.
-The best that can be offered in terms of practical advice is that if the possibilty exists to algorithmically rid a DS model of biases relevant to its application, then this should be done.
+The best that can be offered in terms of practical advice is that if the possibility exists to algorithmically rid a DS model of biases relevant to its application, then this should be done.
 \section{Conclusion}\hypertarget{sec5}{}
-The preceding discussion seems to warrant but two clear take-aways for DS.
+Aside from questions regarding fairness and responsibility we are confident that the PAA is an apt set of principles to guide DS.
+Private information, to name just one example, should be handled just as carefully by a distributional semanticist as by a medical doctor.
+
+The preceding discussion seems to warrant heightened attention for the two remaining points in the PAA.
 First, if a semantic space does contain bias and will be further used in some application, then efforts should be undertaken to debias it.
-Second, if no the DS model cannot be debiased, then it should not be employed externally without further efforts to ensure that it does not discriminate.
-The responsibility for these two points ultimately lie with the designer of the application.
+Second, if the DS model cannot be debiased, then it should not be employed externally without further efforts to ensure that it does not discriminate.
+The responsibility for these two points ultimately lies with the designer of the application.
+Excuses to the effect that the model just mirrors societal bias are unacceptable.
+
+What remains to be done, then, is to constantly re-assess how DS systems are in fact affecting different demographic groups.
+As it is further investigated how DS representations of meaning and societal bias are related to each other, we might be able to elaborate more concrete tests and guidelines for the working distributional semanticist to check her own work for discriminatory effects.
+Perhaps it will also be possible to connect an ethics of DS more systematically with an ethics of NLP in general.
+Until then, the responsible person working on DS will have to assess the impact of her work on a case by case basis.
+
 \section{References}\hypertarget{sec6}{ }

### 2017-06-05 pds/Paper/paper_pds.tex
dis the end

@@
 \usepackage{csquotes}
+\fontsize{10}{12}
+\usepackage[urw-garamond]{mathdesign}
+\usepackage[margin=5cm]{geometry}
 \begin{document}
@@
 
-As recently as 2016, however, a discussion has opened up in a hitherto completely undiscussed domain.
+As recently as 2016, however, discussion has opened up in a hitherto completely undiscussed domain.
 ML techniques used for the computer-based study of language meaning - a field known as \emph{computational distributional semantics} - has come under scrutiny.
@@
 This form of machine bias in word meaning can affect people in discriminatory ways.
-It is quite conceivable that in the near future many applications, for example automated customer services, will operate with automatically generated representations of natural language meaning.
+It is quite conceivable that ian the near future many applications, for example automated customer services, will operate with automatically generated representations of natural language meaning.
 If semantic bias translates into discriminatory decisions being taken, then there is a problem.
@@
 Currently, we want to focus on aspects of discrimination and bias in the outcome of DS.
-This is for two reasons: For one, these are the least explored ethical issues for DS.
-Furthermore, some of the ideas, guidelines and principles concerning privacy, abuse of otherwise legitimate technology ot  or economic impact are directly transferable from the general ML debate to any sub-field that similarly employs big data and algorithms.
+These are the least explored ethical issues for DS.
+Furthermore, some of the ideas, guidelines and principles concerning privacy, abuse of otherwise legitimate technology or economic impact are directly transferable from the general ML debate to any sub-field that similarly employs big data and algorithms.
 Computational DS is no exception.
 Debates concerning privacy etc.
-can follow the same lines as the discussion to establish general purpose ethical principles for ML.
+can follow the same general purpose ethical principles for ML.
 The issue of discrimination and bias, however, has a special relation to DS because the field occupies a unique position of computational language research.
@@
 
-In many cases of undesirable effects that come with these technologies, legislation is behind in adequately protecting people negatively affected by it.
+\newpage In many cases of undesirable effects that come with these technologies, legislation is behind in adequately protecting people negatively affected by it.
 Given this state of affairs, it is important to hold the people who develop or implement such algorithmic systems accountable to the public. The FATML concisely write in their PAA\cite{principles}: 
@@
 
-Concerning the question of fairness, consider a concrete scenario where the problem arises: A recent study has found, for example, that a model of lexical word meaning trained on Google News articles displays significant bias along binary gender dimensions \cite{bolukbasi2016man}.
+Concerning the question of fairness, consider a concrete scenario where the problem arises: A recent study has found that a model of lexical word meaning trained on Google News articles displays significant bias along binary gender dimensions \cite{bolukbasi2016man}.
 Despite the articles in the corpus being written by professional journalists, the word \emph{woman} bears the same semantic relation to \emph{homemaker} in the meaning space as the word \emph{man} to \emph{computer scientist}.
@@
 These are not mere hypotheticals though, but already very concrete issues.
-It is has been reported that when one queries Google with the name of a person, then Google is more likely to show advertisement related to that person's criminal record if the name is associated with black people \cite{sweeney2013discrimination}.
+It is has been reported that when one queries Google with the name of a person, then advertisement related to that person's criminal record is likelier if the name is associated with black people \cite{sweeney2013discrimination}.
 
 This raises a host of questions concerning the provenance of the bias in the model, the adequate response by the responsible distributional semanticist and, not to mention, the nature of meaning.
-Does one have to live with the fact that the meanings discovered in some corpus systematically contain unfavorable connotations for some group of people?
+Is it necessary that the meanings discovered in some corpus systematically contain unfavorable connotations for some group of people?
 After all, the corpus was produced by a particular society, so it should be expected that it mirror that society's prejudices.
@@
 After all, the corpus accurately reflects the fact that there are more men than women who work in computer science \cite{womenincs}.
-So the biased outcome of the search mirrors societal structure, or so would the recommender's excuse might sound.
+So does the biased outcome of the search only mirror societal structure?
+%, or so would the recommender's excuse sound.
 
@@
 Findings that indicate clear instances of bias in widely used corpora can then be used as an indication where  debiasing should have to take place before the DS is put to further use.
-Thus, it appears that a two-fold effort is required in to prevent damage from happening in the employment of algorithmically generated meanings.
-On the one hand, one will constantly have to scan the semantic spaces gathered from corpora for undesired connotations or other semantic relationships within it; on the other, once such biases are certified, efforts need to be undertaken, so that systems using this semantic information do not systematically disadvantage people.
+Thus, it appears that a two-fold effort is required: On the one hand, one will constantly have to scan the semantic spaces gathered from corpora for undesired connotations or other semantic relationships within it; on the other, once such biases are certified, efforts need to be undertaken, so that systems using this semantic information do not systematically disadvantage people.
 
